### 2019-08-30 tex/crossing_over.tex
Re-do all math.

\documentclass{article}

\usepackage{arxiv}
\usepackage{amsmath}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}

\title{A hyperdimensional sequence code inspired by chromosomal crossover}

\author{
  Rich ~Pang\thanks{https://rkp.science} \\
  Computational Neuroscience Center\\
  Department of Physiology and Biophysics\\
  University of Washington\\
  Seattle, WA 98195-7290 \\
  \texttt{rpang at uw dot edu} \\
}

\begin{document}
\maketitle

\begin{abstract}
I present a method for encoding structured information in fixed-length vectors. Drawing on ideas from high-dimensional (HD) computing, the presented scheme encodes elements of information as HD random vectors, and relationships among them (like sequences or trees) by (pseudo-)randomly weaving these vectors together via an operation analogous to chromosomal crossover. Unlike previous HD coding schemes, "crossover" codes retain low Hamming distance to the multi-level sub-relationships that comprise them--all the way down to their base elements--yet high distance to alternative elements or relationships. Accurate decoding can thus occur via greedy reconstruction of the representation. Unlike most machine-learning approaches, crossover (as in other HD computing schemes) requires no training, making it highly amenable to mathematical analysis. I derive several of its formal properties and discuss its relevance to neuroscience and artificial intelligence.
\end{abstract}

% keywords can be removed
\keywords{Hyperdimensional computing \and Vector symbolic architectures \and Sequence encoding \and k-ary tree encoding \and Computational neuroscience}

\section{Introduction}
A core problem in neuroscience is how the brain represents a complex event like a spoken sentence using relatively fixed dimensions. Ideally similar events should have similar representations, and when a new event occurs pre-existing representations shouldn't require substantial rearrangement to prevent confusion.

One approach to solving this has been to assume representations are built from high-dimensional (HD) random vectors \ref{}. Such approach is fundamentally motivated by the observation that independent HD random vectors are very close to orthogonal (and far apart under other distance metrics) with very high probability. Thus, if one considers a codebook where $M$ objects are assigned to $M$ independently sampled HD random vectors, the codes are very unlikely to be confused.

\section{Algorithm}

Consider a system $X$ of $N$ components $x_1, ..., x_N$, each of which can take any of $Z$ states. That is, $x_j \in G$ where $|G| = Z$. Label the states of $G$ as $1,...,Z$ for conciseness. Let $d(X^A, X^B)$ be the Hamming distance between $X^A$ and $X^B$, scaled to range from 0 and 1, and their similarity $s(X^A, X^B) \equiv 1 - d(X^A, X^B)$.

We wish to establish a decodable mapping from arbitrary objects of cognition to system states, such that objects that are intuitively similar to us map to codes with high similarities. We shall start by considering sequences of symbols then generalize to arbitrary finite nested sets (or equivalently k-ary trees) with symbols as elements (leaves).

\subsection{Algorithm}

Given a dictionary $D$ of $M$ symbols $s^1, ..., s^M$, we will first sample a "C-basis" (crossover basis) $B$ and an auxiliary mask function $\mu$ from $P(B)$ and $P(\mu)$, respectively. Using $B$ and $\mu$, we will then deterministically build codes $\{X^Y\}$ of arbitrary sequences $\{Y = (y_1, ..., y_L)\}$, where $y_t \in D$.

\textbf{Setup}

1. Sample a C-basis $B \equiv (X^1, ..., X^M, X^*)$, where each $X^i$ (encoding symbol $s^i$) has components $x^i_j$ sampled uniformly and i.i.d. from $G$, and where $X^*$ is "start" code, also comprised of uniform, i.i.d. components from $G$.

2. Sample a "mask function" $\mu \sim P(\mu)$, where $\mu:(X^A, X^B) \rightarrow [0, 1]^N$ and is represented by a matrix with $N$ rows and $Z^{2N}$ columns, with i.i.d. real-valued elements from $\textrm{Uniform}(0, 1)$. We reference elements of $\mu$ with the notation $\mu(j, X^A, X^B)$.

In practice $\mu$ need not be sampled explicitly, which would take astronomical time; it suffices to use, say, to map $(X^A, X^B)$ using a hash function like SHA-1 to an integer, then sample $N$ times from $\textrm{Uniform}(0, 1)$ from a random number generator seeded with this integer.

\textbf{Encoding}

...

\section{Properties}

\subsection{Sequence codes have i.i.d. components}
\label{sec:iid-comp}

\textbf{Result:}

$$P(X^Y|Y) = \prod\limits_{j=1}^N P(x^Y_j|Y) \quad \textrm{and} \quad P(x^Y_j|Y) = P(x^Y_{j'}|Y) \quad \forall j, j'$$

??? A corollary is that if $Y \sim P(Y)$ then $P(x^Y_j) = P(x^Y_{j'}) \quad \forall j, j' \in [1, ..., N]^2$, and similarly $P(X^Y) $

\textbf{Proof:}

Denoting $Y_{1:t} \equiv (y_1, ..., y_t)$ with $Y_{1:L} = Y$, and $B_j \equiv (x^1_j, ..., x^M_j, x^*_j)$, we want to show that $x_j^{Y_{1:t}}$ are independent across $j$ for all $t$.

We will frequently use the fact that if $A_1, ..., A_N$ are independent, then so are $f(A_1), ..., f(A_N)$. In particular, we note that $x_j^{Y_{1:t}}$ is a strict function of $\xi^t_j \equiv \{x_j^{Y_{1:t-1}}, ..., x_j^{Y_{1:0}}, B_j, \mu(j, X^*, X^{y_1}), \mu(j, X^{Y_{1:1}}, X^{y_2}), ..., \mu(j, X^{Y_{1:t-1}}, X^{y_t})\}.$ Therefore if $\xi^t_j$ are independent across $j$, so are $x_j^{Y_{1:t}}$. Thus, it suffices to show $\xi^t_j$ are independent across $j$ for all $t$.

We proceed by induction:

\textbf{Base case:} $\xi^1_j = \{x^{Y_{1:0}}_j, B_j, \mu(j, X^{Y_{1:0}}, X^{y_1})\} = \{x^*_j, B_j, \mu(j, X^*, X^{y_1})\}$, which are all i.i.d., so $\xi^1_j$ are independent across $j$.

\textbf{Inductive step:} If $\xi^t_j$ are independent across $j$ then so are $\{x_j^{Y_{1:t}}\} \cup \xi^t_j$, since this is just a function of $\xi^t_j$. Further, $\{x_j^{Y_{1:t}}\} \cup \xi^t_j \cup \{\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})\}$ must also be independent across $j$. To see this, consider the following two cases.

First, most likely is that $\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})$ has never been used in the construction of $X^{Y_{1:t-1}}$, so it is independent from $\{x_j^{Y_{1:t}}\} \cup \xi^t_j$ while also independent across $j$. Therefore, their union is independent across $j$.

In the second, less likely case, $\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})$ \textit{was} used to construct $X^{Y_{1:t-1}}$ at some point in the past, i.e. $(X^{Y_{1:t}}, X^{y_{t+1}}) = (X^{Y_{1:t'}}, X^{y_{t'+1}})$ for some $t' < t$. In this case, however, the term $\mu(j, X^{Y_{1:t}}, X^{y_{t+1}}) = \mu(j, X^{Y_{1:t'}}, X^{y_{t'+1}})$ is already included in $\xi^t_j$, and so the independence across $j$ is not affected because no new random variables are added. Thus, $\{x_j^{Y_{1:t}}\} \cup \xi^t_j \cup \{\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})\}$ is independent across $j$.

Since $\{x_j^{Y_{1:t}}\} \cup \xi^t_j \cup \{\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})\} = \xi^{t+1}_j$, we have that $\xi^t_j$ being independent across $j$ implies that $\xi^{t+1}_j$ is independent across $j$, completing the inductive step.

Therefore $x_j^{Y_{1:t}}$ is independent across $j$ for all $t$, in other words $P(X^Y|Y) = \prod\limits_{j=1}^N P(x^Y_j|Y)$.

The equality $P(x^Y_j|Y) = P(x^Y_{j'}|Y) \quad \forall j, j'$ arises because no component is treated differently from any other in our algorithm, so symmetry implies they can have no statistical differences.

\subsection{Component distributions are uniform}
\label{sec:prop-eqty-comp-dstr}

\textbf{Result:}

$$P(x^Y_j = g|Y) = P(x^Y_j = g'|Y) = \frac{1}{Z} \quad \forall g, g' \in G \quad \textrm{so} \quad P(X^Y|Y) = \prod\limits_{j=1}^N P(x^Y_j|Y) = \frac{1}{Z^N}$$

This implies neither $P(x^Y_j = g|Y)$ nor $P(X^Y|Y)$ are a function of $Y$. Thus, if $Y \sim P(Y)$ we then also have

$$P(x_j) = \sum\limits_Y P(x^Y_j|Y)P(Y) = \frac{1}{Z} \quad \textrm{and} \quad P(X) = \sum\limits_Y P(X^Y|Y)P(Y) = \frac{1}{Z^N}.$$

\textbf{Proof:}

$P(x^Y_j = g|Y) = P(x^Y_j = g'|Y)$ for $g \neq g'$ arises because all $g$ in our model are treated identically, so symmetry implies any functions of them can have no statistical differences. Since these must add to $1$ when summed over $g \in G$, and $|G| = Z$, we must also have $P(x^Y_j = g|Y) = 1/Z.$

\subsection{Code statistics are invariant to sequence statistics}

\textbf{Result:}

$$P(X^Y|Y) \textrm{ does not depend on any statistic (e.g. length, symbol repetitions, etc) of } Y.$$

\textbf{Proof:}

This follows from \ref{sec:prop-eqty-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$. Therefore the statistics of crossover codes do not depend on any aspect about the sequence being encoded. This distinguishes crossover, for instance, from representations that grow more or less crowded with the complexity of the sequence they represent.

\subsection{Information about sequence identity is evenly distributed across the code}
\label{sec:perf-info-dstr}

Given $B$, $\mu$, and a distribution over sequences $P_Y(Y)$ with entropy $H[Y]$, define the information $(I_j | B, \mu)$ in component $j$ as the mutual information between $Y$ and $x_j$, i.e.

$$I_j | B, \mu \equiv MI[Y|x_j; B, \mu] = H[Y] - H[Y|x_j; B, \mu] = H[Y] - \sum_g H[Y|x_j = g; B, \mu] P_Y(x_j = g| B, \mu)$$

\textbf{Result:}

$$P(I_1, ..., I_N) = \prod_j P(I_j) \quad \textrm{and} \quad P(I_j) = P(I_{j'}) \quad \forall j, j'.$$

In other words, each component contains independent and equal information about the sequence. (Note that this does not imply $I_j = I_{j'}$ for any given $B$ and $\mu$, since there are $B, \mu$ in which some $x_j$ are less informative about $Y$ than others, e.g. if some $x_j$ by chance has the same c-state for all $Y$.) This means that the code's robustness to random and targeted lesions are equivalent, depending only on how many components were damaged.

\textbf{Proof:}

This follows from \ref{sec:iid-comp}, since the components for one sequence $Y$ are i.i.d. Therefore, while the $x^Y_j$ and $x^{Y'}_j$ might be dependent (e.g. if $Y$ and $Y'$ are very similar; see below), there can be no dependencies between $x^Y_j$ and $x^{Y'}_{j'}$ for $j \neq j'$. Therefore $\{x^{Y_1}_j, ..., x^{Y_{N_Y}}_j\}$ is independent across $j$. Since $I_j$ is a function of $\{x^{Y_1}_j, ..., x^{Y_{N_Y}}_j\}$, $I_j$ is independent across $j$ also. Since no $j$ is treated specially in our algorithm, $I_j$ must also be identically distributed, so given $P(Y)$, $I_j$ is i.i.d.

\subsection{Similarity between sequences}

\textbf{Result:}

Since $x^Y_j$ are i.i.d. across $j$ (\ref{sec:iid-comp}) it follows that $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$. Thus, $\textrm{E}[s(X^Y, X^{Y'})] = q$ and $\textrm{Var}[s(X^Y, X^{Y'})] = q(1-q)/N$, and it suffices to only compute $q$. For conciseness, we drop the $j$ and write $x \equiv x_j$ from now on.

\textbf{Proof:}

\textbf{Prelude:}

We first define and compute a few useful quantities that will simplify the rest of our calculations.

1. $P(x^Y \leftarrow y_t)$ is the probability that $x^Y$ was taken from the $t$-th element of $Y$, that is, $1/(t+1)^\gamma$ times the probability it was not overwritten between $t+1$ and $L$ (including occasions where it was overwritten by the same state). For notational convenience we call this $f^{\gamma,L}_t$. We then have:

$$f^{\gamma,L}_t \equiv P(x^Y \leftarrow y_t) = \frac{1}{(t+1)^\gamma}\prod\limits_{t' = t+1}^L\left(1 - \frac{1}{(t'+1)^\gamma}\right).$$

Note that this is not a function of $x^Y$. Also note that $f^{1,L}_t = 1/(L+1)$, which is not a function of $t$.

2. $P(x^Y \leftarrow i)$ is the probability that $x^Y$ was taken from symbol $i$, which will depend on how many times $i$ was present in $Y$. We compute this by summing the probabilities of all ways this could have occurred.

Suppose $y_t = i$ for $t_1 < ... < t_{n^i}$, with $\sum_i n^i = L$. We can divide the combinatorial ways $x^Y$ could have ended up being taken from $i$ into $n^i$ cases. First, $x^Y$ could have been taken from $y_{t_1}$ then never overwritten, whose probability is $f^{\gamma,L}_{t_1}$. Next, it could have been taken from $y_{t_2}$ then never overwritten (regardless of whether it was taken from $y_{t_1}$), whose probability is $f^{\gamma,L}_{t_2}$. And so forth until $t_{n^i}$. Note that these $n^i$ cases are mutually non-overlapping and cover all possible ways that $x^Y$ could have been taken from $i$. Thus

$$P(x^Y \leftarrow i) = f^{\gamma,L}_{t_1} + ... + f^{\gamma,L}_{t_{n^i}}$$.

When $\gamma = 1$, $P(x^Y \leftarrow i) = n/(L+1)$.

3. $P(x^Y \leftarrow Y_{1:t})$ is the probability that $x^Y$ was not overwritten from $t+1$ to $L$. This is just

$$P(x^Y \leftarrow Y_{1:t}) = \prod\limits_{t' = t+1}^L \left(1 - \frac{1}{(t'+1)^\gamma}\right) = (t+1)^\gamma f^{\gamma,L}_t.$$

When $\gamma = 1$, $P(x^Y \leftarrow Y_{1:t}) = (t+1)/(L+1)$.

4. $P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t})$ is the probability that $x^Y$ came from $i$ given that it did not come from $Y_{1:t}$ (i.e. given that it was overwritten at some point from $t+1$ to $L$). Suppose $y_s = i$ for $t < s_1 < ... < s_{n^i_v}$ where $n^i_v$ is the number of times symbol $i$ appears in $Y_{t+1:L} \equiv (y_{t+1}, ..., y_L)$, with $\sum\limits_i n^i_v = L-t$.

We compute $P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t})$ by summing over the probabilities of all the ways this could have occurred. Similar to the calculation for $P(x^Y \leftarrow i)$, we have

$$P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t}) = P(x^Y \leftarrow y_{s_1}| x^Y \not \leftarrow Y_{1:t}) + ... + P(x^Y \leftarrow y_{s_{n^i_v}}| x^Y \not \leftarrow Y_{1:t}).$$

By Bayes' Rule:

$$P(x^Y \leftarrow y_{s_k}| x^Y \not \leftarrow Y_{1:t}) = \frac{P(x^Y \not \leftarrow Y_{1:t}| x^Y \leftarrow y_{s_k})P(x^Y \leftarrow y_{s_k})}{P(x^Y \not\leftarrow Y_{1:t})} = \frac{P(x^Y \leftarrow y_{s_k})}{P(x^Y \not\leftarrow Y_{1:t})}$$

since if $x^Y$ came from $y_{s_k}$, which is in the part of $Y$ after $t$, then $x^Y$ could have not have come from $Y_{1:t}$. We have already computed the numerator and denominator (which is just $1 - P(x^Y \leftarrow Y_{1:t}))$, so

$$P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t}) = \frac{1}{P(x^Y \not\leftarrow Y_{1:t})}\left( f^{\gamma,L}_{s_1} + ... + f^{\gamma,L}_{s_{n^i_v}} \right).$$

When $\gamma = 1$ this reduces to $P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t}) = \frac{n^i_v}{L-t}$, i.e. it is proportional to how many times $i$ appears in the second part of the sequence $Y_{t+1:L}$.

We now calculate how similar a sequence code is to any symbol code, as well as to other sequence codes.

\textbf{Similarity between symbol and sequence codes:}

If $y_t = i$ for $0 < t_1 < ... < t_{n^i}$, then

$$P(x^Y = x^i) = \frac{1}{Z} + \sum\limits_{t \in (t_1, ..., t_{n^i})}\left(1 - \frac{1}{Z}\right)f^{\gamma,L}_t.$$

For $\gamma = 1$:

$$P(x^Y = x^i) = \frac{n}{L+1} + \frac{1}{Z}\frac{L+1-n}{L+1}$$

which tends to $n/(L+1)$ for large $Z$.

\textbf{Proof}:

$$P(x^Y = x^i) = P(x^Y \leftarrow i)P(x^Y = x^i|x^Y \leftarrow i) + P(x^Y \not \leftarrow i)P(x^Y = x^i|x^Y \not \leftarrow i)$$

$$= P(x^Y \leftarrow i) + P(x^Y \not \leftarrow i)\frac{1}{Z}$$

$$= \sum\limits_{t \in (t_1, ..., t_{n^i})}f^{\gamma,L}_t + \left(1 - \sum\limits_{t \in (t_1, ..., t_{n^i})}f^{\gamma,L}_t\right)\frac{1}{Z}
= \sum\limits_{t \in (t_1, ..., t_{n^i})}\left(1 - \frac{1}{Z}\right)f^{\gamma,L}_t + \frac{1}{Z}$$

\textbf{Similarity between codes for two sequences:}

Let $Y = (y_1, ..., y_L)$ and $Y' = (y'_1, .., y'_{L'})$ be two sequences. Suppose they share a starting subsequence of length $t$, i.e. $y_1 = y'_1, ..., y_t = y'_t$, but either $y_{t+1} \neq y'_{t+1}$, $t = L$, or $t = L'$. We write this condition as $Y_{1:t} = Y'_{1:t}$. Similarly, we refer to the second parts of the sequences (the subsequences in which they differ) as $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively.

Let $u^i$ index the times $i$ appears in $Y_{1:t}$, $v^i$ the times $i$ appears in $Y_{t+1:L}$, and $(v')^i$ the times $i$ appears in $Y'_{t+1:L'}$. Let $n^i_u$ count how many times $i$ appears in $Y_{1:t}$, $n^i_v$ the number of times $i$ appears in $Y_{t+1:L}$, and $n^i_{v'}$ the number of times $i$ appears in $Y'_{t+1:L'}$. We also write these in vector format as $\mathbf{n}_u$, $\mathbf{n}_v$, and $\mathbf{n}_{v'}$, respectively, where each is a vector of length $M$.

Note that we must deal specially with starting subsequences because sequence codes are constructed from beginning to end.

Then we have:

Wrong: Note that when $Y$ and $Y'$ do not start with the same subsequence (i.e. $t = 0$) this reduces to

$$P(x^Y = x^{Y'}) = \frac{1}{(L+1)(L'+1)}\left[1 + \frac{L + L' + LL'}{Z} + \frac{Z-1}{Z}\mathbf{n}^T\mathbf{n}'\right]$$

\textbf{Proof:}

Since $Y_{1:t} = Y'_{1:t}$,  first note that $X^{Y_{1:t}} = X^{Y'_{1:t}}$, so $x^{Y_{1:t}} = x^{Y'_{1:t}}$. We wish to find $P(x^Y = x^{Y'})$.

The event $x^Y = x^{Y'}$ can occur in 4 distinct ways, whose probabilities sum to $P(x^Y = x^{Y'})$. At present we work under the assumption that all symbol codes are different and that $Y \neq Y' \implies X^{Y} \neq X^{Y'}.$

\textbf{Case 1}: $(x^Y \leftarrow Y_{1:t}) \land (x^{Y'} \leftarrow Y'_{1:t})$, i.e. neither $x$ is overwritten at $t+1$ or later. These events are independent, so

$$P(\textrm{Case 1}) = P(x^Y \leftarrow Y_{1:t})P(x^{Y'} \leftarrow Y'_{1:t}) = (t+1)^\gamma f^{\gamma,L}_t(t+1)^\gamma f^{\gamma,L'}_t$$

When $\gamma = 1$, then:

$$P(\textrm{Case 1}) = \frac{t+1}{L+1}\frac{t+1}{L'+1}$$.

\textbf{Case 2}: $(x^Y \not\leftarrow Y_{1:t}) \land (x^{Y'} \not\leftarrow Y'_{1:t}) \land (x^Y = x^{Y'})$, i.e. $x$ is overwritten in both sequences at some point between $t+1$ and $L$ or $L'$, but by the same c-state. Since the first two conditions are again independent we have

$$P(\textrm{Case 2}) = P(x^Y \not\leftarrow Y_{1:t})P(x^{Y'} \not\leftarrow Y'_{1:t})P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$$

$$= [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t]P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})

where we recall that $Y_{1:t} = Y'_{1:t}$. When $\gamma = 1$ this simplifies to

$$P(\textrm{Case 2}) = \frac{L-t}{L+1}\frac{L'-t}{L'+1}P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$$

To find $P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$ we note that if both $x$'s are overwritten, they can end up with the same c-state if (1) $x^Y$ and $x^{Y'}$ are taken from the same symbol $i$, whose probability we write as $P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$ or (2) $x^Y$ and $x^{Y'}$ are taken from different symbol codes but end up with the same c-state. That is,

$$P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) =$$

$$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) + [1 - P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow Y_{1:t})]\frac{1}{Z}.$$

$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}})$ is the sum over $i \in \{1, ..., M\}$ of the probabilities that both $x^Y$ and $x^{Y'}$ ended up being taken from $i$. These events are independent, so:

$$P(x^Y, x^{Y'} \leftarrow i|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = P(x^Y \leftarrow i|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}})P(x^{Y'} \leftarrow i|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}})$$

$$ = \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}.$$

When $\gamma = 1$, this simplifies to

$$\frac{n^i_v}{L-t}\frac{n^i_{v'}}{L'-t}$$.

Thus

$$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}.$$

When $\gamma = 1$ this is

$$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = \sum_i \frac{n^i_vn^i_{v'}}{(L-t)(L'-t)} = \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)}$$

i.e. when $\gamma = 1$ the probability that $x^Y$ came from the same symbol, given that it was overwritten in the second part of both sequence constructions, is just the normalized dot product of the symbol-count vectors for the second parts of the two sequences.

Thus

$$P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = $$

$$\sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} + \left(1 - \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}$$

When $\gamma = 1$

$$P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}.$$

Thus

$$P(\textrm{Case 2}) = [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t] \times$$

$$\left[
\sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} + \left(1 - \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
\right]$$

When $\gamma = 1$:

$$P(\textrm{Case 2}) = \frac{L-t}{L+1}\frac{L'-t}{L'+1}\left[\frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}\right].$$

\textbf{Case 3}: $(x^Y \leftarrow Y_{1:t}) \land (x^{Y'} \not\leftarrow Y'_{1:t}) \land (x^Y = x^{Y'})$, i.e. $x$ is untouched in the first sequence but overwritten in the second, but happens to be overwritten by $x^{Y_{1:t}}$. Once again the first two events are independent, so

$$P(\textrm{Case 3}) = (t+1)^\gamma f^{\gamma,L}_t[1-(t+1)^\gamma f^{\gamma,L'}_t]P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}).$$

When $\gamma = 1$:

$$P(\textrm{Case 3}) = \frac{t+1}{L+1}\frac{L'-t}{L'+1}P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}).$$

By similar reasoning as in Case 2, the event $(x^Y = x^{Y'}|x^Y \leftarrow X^{Y_{1:t}}, x^{Y'} \not\leftarrow X^{Y'_{1:t}})$ can occur if (1) $x^Y$ and $x^{Y'}$ are taken from the same symbol $i$ or (2) $x^Y$ and $x^{Y'}$ are taken from different symbols but end up with the same c-state.

Following similar logic as in Case 2, except swapping $Y^{t+1:L}$ with $Y^{1:t}$ we thus have

$$P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}) =$$

$$\left[
\sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} 
+ \left(1 - \sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
\right]$$

When $\gamma = 1$:

$$P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}) =
\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}$$

Thus,

$$P(\textrm{Case 3}) = (t+1)^\gamma f^{\gamma,L}_t[1-(t+1)^\gamma f^{\gamma,L'}_t] \times 
\left[
\sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} 
+ \left(1 - \sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
\right]$$

And when $\gamma = 1$:

$$P(\textrm{Case 3}) = \frac{t+1}{L+1}\frac{L'-t}{L'+1}\left[
\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}
\right].$$

\textbf{Case 4}: $(x^Y \not \leftarrow Y_{1:t}) \land (x^{Y'} \leftarrow Y'_{1:t}) \land (x^Y = x^{Y'})$, i.e. $x$ is untouched in the second sequence but overwritten in the first, but happens to be overwritten by $x^{Y'_{1:t}}$. This is symmetric to Case 3, so

$$P(\textrm{Case 4}) = [1 - (t+1)^\gamma f^{\gamma,L}_t](t+1)^\gamma f^{\gamma,L'}_t \times \left[
\sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t} 
+ \left(1 - \sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t}\right)\frac{1}{Z}
\right]$$

And when $\gamma = 1$:

$$P(\textrm{Case 4}) = \frac{L'-t}{L+1}\frac{t+1}{L'+1}\left[
\frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} + \left(1 - \frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} \right)\frac{1}{Z}
\right].$$

\textbf{Sum over all cases}

$$P(x^Y = x^{Y'}) = P(\textrm{Case 1}) + P(\textrm{Case 2}) + P(\textrm{Case 3}) + P(\textrm{Case 4}) = $$

$$
(t+1)^\gamma f^{\gamma,L}_t(t+1)^\gamma f^{\gamma,L'}_t + [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t] \times$$

$$\left[
\sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} + \left(1 - \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
\right]$$

$$+(t+1)^\gamma f^{\gamma,L}_t[1-(t+1)^\gamma f^{\gamma,L'}_t] \times 
\left[
\sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} 
+ \left(1 - \sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
\right]
$$

$$+[1 - (t+1)^\gamma f^{\gamma,L}_t](t+1)^\gamma f^{\gamma,L'}_t \times \left[
\sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t} 
+ \left(1 - \sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t}\right)\frac{1}{Z}
\right]
$$

When $\gamma = 1$, then $P(x^Y = x^{Y'}) = $

$$\frac{t+1}{L+1}\frac{t+1}{L'+1}
+ \frac{L-t}{L+1}\frac{L'-t}{L'+1}\left[\frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}\right]
$$

$$+\frac{t+1}{L+1}\frac{L'-t}{L'+1}\left[
\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}
\right]
$$

$$+\frac{L'-t}{L+1}\frac{t+1}{L'+1}\left[
\frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} + \left(1 - \frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} \right)\frac{1}{Z}
\right]
$$

\textbf{OLD STARTS HERE}

This means that when $Y'$ is a starting subsequence of $Y$, i.e. $Y = (y'_1, ..., y'_t, y_{t+1}, ..., y_L)$ then the two sequences' expected similarity is equal to

$$P(x^{Y'} = x^Y) = \frac{1}{(L+1)(t+1)}\left((t+1)^2 + \frac{Z-1}{Z}\mathbf{n}^T_{1:t}\mathbf{n}_{t+1:L} + \frac{(t+1)(L-t)}{Z}\right).$$

\section{Comparison to existing hyperdimensional computing codes}

\bibliographystyle{unsrt}  
%\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
%%% and comment out the ``thebibliography'' section.

%%% Comment out this section when you \bibliography{references} is enabled.
\begin{thebibliography}{1}

\bibitem{kour2014real}
George Kour and Raid Saabne.
\newblock Real-time segmentation of on-line handwritten arabic script.
\newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
  International Conference on}, pages 417--422. IEEE, 2014.

\bibitem{kour2014fast}
George Kour and Raid Saabne.
\newblock Fast classification of handwritten on-line arabic characters.
\newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
  International Conference of}, pages 312--318. IEEE, 2014.

\bibitem{hadash2018estimate}
Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
  Jacovi.
\newblock Estimate and replace: A novel approach to integrating deep neural
  networks with existing applications.
\newblock {\em arXiv preprint arXiv:1804.09028}, 2018.

\end{thebibliography}

\section{Appendix}

Blah blah

\end{document}

### 2019-09-01 ms/crossing_over.tex
Rename.

--- 2019-08-30 tex/crossing_over.tex
@@
 I present a method for encoding structured information in fixed-length vectors.
-Drawing on ideas from high-dimensional (HD) computing, the presented scheme encodes elements of information as HD random vectors, and relationships among them (like sequences or trees) by (pseudo-)randomly weaving these vectors together via an operation analogous to chromosomal crossover.
-Unlike previous HD coding schemes, "crossover" codes retain low Hamming distance to the multi-level sub-relationships that comprise them--all the way down to their base elements--yet high distance to alternative elements or relationships.
+Drawing on ideas from high-dimensional (HD) computing, the presented scheme encodes elements of information as HD random vectors and sequences of elements by (pseudo-)randomly weaving these vectors together via an operation analogous to chromosomal crossover.
+"Crossover" codes retain low Hamming distance to both their elements and the codes for the subsequences used to build them, yet high distance to alternative elements or sequences.
 Accurate decoding can thus occur via greedy reconstruction of the representation.
 Unlike most machine-learning approaches, crossover (as in other HD computing schemes) requires no training, making it highly amenable to mathematical analysis.
-I derive several of its formal properties and discuss its relevance to neuroscience and artificial intelligence.
+I derive several properties of crossover and discuss its relevance to neuroscience and artificial intelligence.
 \end{abstract}
@@
 % keywords can be removed
-\keywords{Hyperdimensional computing \and Vector symbolic architectures \and Sequence encoding \and k-ary tree encoding \and Computational neuroscience}
+\keywords{Hyperdimensional computing \and Vector symbolic architectures \and Sequence encoding \and Computational neuroscience \and Artificial intelligence \and Cognitive computing}
 
 \section{Introduction}
-A core problem in neuroscience is how the brain represents a complex event like a spoken sentence using relatively fixed dimensions.
+A core problem in neuroscience and artificial intelligence is how to represent a complex event like a spoken sentence using relatively fixed dimensions.
 Ideally similar events should have similar representations, and when a new event occurs pre-existing representations shouldn't require substantial rearrangement to prevent confusion.
@@
 \subsection{Sequence codes have i.i.d. components}
-\label{sec:iid-comp}
-
-\textbf{Result:}
+\label{sec:prop-comp-iid}
 
@@
 
-??? A corollary is that if $Y \sim P(Y)$ then $P(x^Y_j) = P(x^Y_{j'}) \quad \forall j, j' \in [1, ..., N]^2$, and similarly $P(X^Y) $
+\subsection{Component distributions are uniform over c-states}
+\label{sec:prop-unif-comp-dstr}
 
-\textbf{Proof:}
+$$P(x^Y_j = g|Y) = P(x^Y_j = g'|Y) = \frac{1}{Z} \quad \forall g, g' \in G \quad \textrm{so} \quad P(X^Y|Y) = \prod\limits_{j=1}^N P(x^Y_j|Y) = \frac{1}{Z^N}$$
+
+This implies neither $P(x^Y_j = g|Y)$ nor $P(X^Y|Y)$ are a function of $Y$. Thus, if $Y \sim P(Y)$ we then also have
+
+$$P(x_j) = \sum\limits_Y P(x^Y_j|Y)P(Y) = \frac{1}{Z} \quad \textrm{and} \quad P(X) = \sum\limits_Y P(X^Y|Y)P(Y) = \frac{1}{Z^N}.$$
+
+\subsection{Code statistics are invariant to sequence statistics}
+
+$$P(X^Y|Y) \textrm{ does not depend on any statistic (e.g. length, symbol repetitions, etc) of } Y.$$
+
+\subsection{Sequence information is i.i.d. across components}
+\label{sec:prop-info-iid}
+
+Given $B$, $\mu$, and a distribution over sequences $P_Y(Y)$ with entropy $H[Y]$, define the information $I_j | B, \mu$ in component $j$ as the mutual information between $Y$ and $x_j$, i.e.
+
+$$I_j | B, \mu \equiv MI[Y|x_j; B, \mu] = H[Y] - H[Y|x_j; B, \mu] = H[Y] - \sum_g H[Y|x_j = g; B, \mu] P_Y(x_j = g| B, \mu).$$
+
+Then:
+
+$$P(I_1, ..., I_N) = \prod_j P(I_j) \quad \textrm{and} \quad P(I_j) = P(I_{j'}) \quad \forall j, j'.$$
+
+In other words, each component contains independent and equal information about the sequence.
+(Note that this does not imply $I_j = I_{j'}$ for any given $B$ and $\mu$, since there are $B, \mu$ in which some $x_j$ are less informative about $Y$ than others, e.g.
+if some $x_j$ by chance has the same c-state for all $Y$.) This means that the code's robustness to random and targeted lesions are equivalent, depending only on how many components were damaged.
+
+\section{Code similarities}
+
+We give results for $\gamma = 1$ here, with results for arbitrary $\gamma$ in the appendix.
+
+Since $x^Y_j$ are i.i.d.
+across $j$ (\ref{sec:prop-comp-iid}) it follows that $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$.
+Thus, $\textrm{E}[s(X^Y, X^{Y'})] = q$ and $\textrm{Var}[s(X^Y, X^{Y'})] = q(1-q)/N$, and it suffices to only compute $q$.
+For conciseness, we drop the $j$ and write $x \equiv x_j$ from now on.
+
+\subsection{Similarity between symbol and sequence codes}
+
+If there are $n^i$ repetitions of symbol $i$ appears in $Y$, then:
+
+$$P(x^Y = x^i) = \frac{n^i}{L+1} + \frac{1}{Z}\frac{L+1-n^i}{L+1}$$
+
+which tends to $n^i/(L+1)$ for large $Z$.
+
+\subsection{Similarity between codes for two sequences}
+
+Let $Y = (y_1, ..., y_L)$ and $Y' = (y'_1, .., y'_{L'})$ be two sequences that share a starting subsequence of length $t$.
+I.e.
+$y_1 = y'_1, ..., y_t = y'_t$, but either $y_{t+1} \neq y'_{t+1}$, $t = L$, or $t = L'$.
+We write that $Y_{1:t} = Y'_{1:t}$, and we refer to the second parts of the sequences as $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively.
+We must deal specially with starting subsequences because sequence codes are constructed from beginning to end.
+
+Let $\mathbf{n}_u$ be the length-$M$ symbol count vector for $Y_{1:t} = Y'_{1:t}$ with $\mathbf{1}^T\mathbf{n}_u = t$.
+Let $\mathbf{n}_v$ and $\mathbf{n}_{v'}$ be the symbol count vectors for $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively, with $\mathbf{1}^T\mathbf{n}_{v} = L - t$ and $\mathbf{1}^T\mathbf{n}_{v'} = L' - t$.
+Then:
+
+$$P(x^Y = x^{Y'}) = \frac{t+1}{L+1}\frac{t+1}{L'+1}$$
+
+$$+ \frac{1}{L+1}\frac{1}{L'+1}\left[\mathbf{n}^T_v\mathbf{n}_{v'} + \left((L-t)(L'-t) - \mathbf{n}^T_v\mathbf{n}_{v'} \right)\frac{1}{Z}\right]
+$$
+
+$$+\frac{1}{L+1}\frac{1}{L'+1}\left[
+\mathbf{n}^T_{u}\mathbf{n}_{v'} + \left((t+1)(L'-t) - \mathbf{n}^T_{u}\mathbf{n}_{v'} \right)\frac{1}{Z}
+\right]
+$$
+
+$$+\frac{1}{L+1}\frac{1}{L'+1}\left[
+\mathbf{n}^T_{v}\mathbf{n}_{u} + \left((L-t)(t+1) - \mathbf{n}^T_{v}\mathbf{n}_{u} \right)\frac{1}{Z}
+\right]
+$$
+
+This means that when $Y'$ is a starting subsequence of $Y$, i.e. $Y' = Y_{1:t}$ then the two sequences' expected similarity is
+
+$$P(x^{Y'} = x^Y) = \frac{1}{(L+1)(t+1)}\left((t+1)^2 + \frac{Z-1}{Z}\mathbf{n}^T_{1:t}\mathbf{n}_{t+1:L} + \frac{(t+1)(L-t)}{Z}\right).$$
+
+\section{Comparison to existing hyperdimensional computing codes}
+
+\bibliographystyle{unsrt}  
+%\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
+%%% and comment out the ``thebibliography'' section.
+
+%%% Comment out this section when you \bibliography{references} is enabled.
+\begin{thebibliography}{1}
+
+\bibitem{kour2014real}
+George Kour and Raid Saabne.
+\newblock Real-time segmentation of on-line handwritten arabic script.
+\newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
+  International Conference on}, pages 417--422. IEEE, 2014.
+
+\bibitem{kour2014fast}
+George Kour and Raid Saabne.
+\newblock Fast classification of handwritten on-line arabic characters.
+\newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
+  International Conference of}, pages 312--318. IEEE, 2014.
+
+\bibitem{hadash2018estimate}
+Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
+  Jacovi.
+\newblock Estimate and replace: A novel approach to integrating deep neural
+  networks with existing applications.
+\newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
+
+\end{thebibliography}
+
+\section{Appendix}
+
+\subsection{Proof: Sequence codes have i.i.d. components}
 
@@
 
-\subsection{Component distributions are uniform}
-\label{sec:prop-eqty-comp-dstr}
-
-\textbf{Result:}
-
-$$P(x^Y_j = g|Y) = P(x^Y_j = g'|Y) = \frac{1}{Z} \quad \forall g, g' \in G \quad \textrm{so} \quad P(X^Y|Y) = \prod\limits_{j=1}^N P(x^Y_j|Y) = \frac{1}{Z^N}$$
-
-This implies neither $P(x^Y_j = g|Y)$ nor $P(X^Y|Y)$ are a function of $Y$. Thus, if $Y \sim P(Y)$ we then also have
-
-$$P(x_j) = \sum\limits_Y P(x^Y_j|Y)P(Y) = \frac{1}{Z} \quad \textrm{and} \quad P(X) = \sum\limits_Y P(X^Y|Y)P(Y) = \frac{1}{Z^N}.$$
-
-\textbf{Proof:}
+\subsection{Proof: Component distributions are uniform over c-states}
 
@@
 
-\subsection{Code statistics are invariant to sequence statistics}
+\subsection{Proof: Code statistics are invariant to sequence statistics}
 
-\textbf{Result:}
-
-$$P(X^Y|Y) \textrm{ does not depend on any statistic (e.g. length, symbol repetitions, etc) of } Y.$$
-
-\textbf{Proof:}
-
-This follows from \ref{sec:prop-eqty-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$.
+This follows immediately from \ref{sec:prop-unif-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$.
 Therefore the statistics of crossover codes do not depend on any aspect about the sequence being encoded.
@@
 
-\subsection{Information about sequence identity is evenly distributed across the code}
-\label{sec:perf-info-dstr}
+\subsection{Proof: Sequence information is i.i.d. across components}
 
-Given $B$, $\mu$, and a distribution over sequences $P_Y(Y)$ with entropy $H[Y]$, define the information $(I_j | B, \mu)$ in component $j$ as the mutual information between $Y$ and $x_j$, i.e.
-
-$$I_j | B, \mu \equiv MI[Y|x_j; B, \mu] = H[Y] - H[Y|x_j; B, \mu] = H[Y] - \sum_g H[Y|x_j = g; B, \mu] P_Y(x_j = g| B, \mu)$$
-
-\textbf{Result:}
-
-$$P(I_1, ..., I_N) = \prod_j P(I_j) \quad \textrm{and} \quad P(I_j) = P(I_{j'}) \quad \forall j, j'.$$
-
-In other words, each component contains independent and equal information about the sequence.
-(Note that this does not imply $I_j = I_{j'}$ for any given $B$ and $\mu$, since there are $B, \mu$ in which some $x_j$ are less informative about $Y$ than others, e.g.
-if some $x_j$ by chance has the same c-state for all $Y$.) This means that the code's robustness to random and targeted lesions are equivalent, depending only on how many components were damaged.
-
-\textbf{Proof:}
-
-This follows from \ref{sec:iid-comp}, since the components for one sequence $Y$ are i.i.d.
-Therefore, while the $x^Y_j$ and $x^{Y'}_j$ might be dependent (e.g.
+Since the components for one sequence $Y$ are i.i.d.
+(\ref{sec:prop-comp-iid}) Therefore, while the $x^Y_j$ and $x^{Y'}_j$ might be dependent (e.g.
 if $Y$ and $Y'$ are very similar; see below), there can be no dependencies between $x^Y_j$ and $x^{Y'}_{j'}$ for $j \neq j'$.
@@
 
-\subsection{Similarity between sequences}
-
-\textbf{Result:}
+\subsection{Code similarity derivations}
 
 Since $x^Y_j$ are i.i.d.
-across $j$ (\ref{sec:iid-comp}) it follows that $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$.
+across $j$ (\ref{sec:prop-comp-iid}) it follows that $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$.
 Thus, $\textrm{E}[s(X^Y, X^{Y'})] = q$ and $\textrm{Var}[s(X^Y, X^{Y'})] = q(1-q)/N$, and it suffices to only compute $q$.
@@
 
-\textbf{Proof:}
-
-\textbf{Prelude:}
-
-We first define and compute a few useful quantities that will simplify the rest of our calculations.
+We define and compute a few useful quantities that will keep our calculations more precise.
 
@@
 
-\textbf{Similarity between symbol and sequence codes:}
+\subsubsection{Derivation: Similarity between symbol and sequence codes}
 
-If $y_t = i$ for $0 < t_1 < ... < t_{n^i}$, then
-
-$$P(x^Y = x^i) = \frac{1}{Z} + \sum\limits_{t \in (t_1, ..., t_{n^i})}\left(1 - \frac{1}{Z}\right)f^{\gamma,L}_t.$$
-
-For $\gamma = 1$:
-
-$$P(x^Y = x^i) = \frac{n}{L+1} + \frac{1}{Z}\frac{L+1-n}{L+1}$$
-
-which tends to $n/(L+1)$ for large $Z$.
-
-\textbf{Proof}:
-
-$$P(x^Y = x^i) = P(x^Y \leftarrow i)P(x^Y = x^i|x^Y \leftarrow i) + P(x^Y \not \leftarrow i)P(x^Y = x^i|x^Y \not \leftarrow i)$$
-
-$$= P(x^Y \leftarrow i) + P(x^Y \not \leftarrow i)\frac{1}{Z}$$
+$$P(x^Y = x^i) = P(x^Y \leftarrow i)P(x^Y = x^i|x^Y \leftarrow i) + P(x^Y \not \leftarrow i)P(x^Y = x^i|x^Y \not \leftarrow i)
+= P(x^Y \leftarrow i) + P(x^Y \not \leftarrow i)\frac{1}{Z}$$
 
@@
 
-\textbf{Similarity between codes for two sequences:}
-
-Let $Y = (y_1, ..., y_L)$ and $Y' = (y'_1, .., y'_{L'})$ be two sequences.
-Suppose they share a starting subsequence of length $t$, i.e.
-$y_1 = y'_1, ..., y_t = y'_t$, but either $y_{t+1} \neq y'_{t+1}$, $t = L$, or $t = L'$.
-We write this condition as $Y_{1:t} = Y'_{1:t}$.
-Similarly, we refer to the second parts of the sequences (the subsequences in which they differ) as $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively.
+\subsubsection{Derivation: Similarity between codes for two sequences}
 
@@
 Let $n^i_u$ count how many times $i$ appears in $Y_{1:t}$, $n^i_v$ the number of times $i$ appears in $Y_{t+1:L}$, and $n^i_{v'}$ the number of times $i$ appears in $Y'_{t+1:L'}$.
-We also write these in vector format as $\mathbf{n}_u$, $\mathbf{n}_v$, and $\mathbf{n}_{v'}$, respectively, where each is a vector of length $M$.
-
-Note that we must deal specially with starting subsequences because sequence codes are constructed from beginning to end.
-
-Then we have:
-
-Wrong: Note that when $Y$ and $Y'$ do not start with the same subsequence (i.e. $t = 0$) this reduces to
-
-$$P(x^Y = x^{Y'}) = \frac{1}{(L+1)(L'+1)}\left[1 + \frac{L + L' + LL'}{Z} + \frac{Z-1}{Z}\mathbf{n}^T\mathbf{n}'\right]$$
-
-\textbf{Proof:}
+We also write these in vector format as $\mathbf{n}_u$, $\mathbf{n}_v$, and $\mathbf{n}_{v'}$, respectively.
 
@@
 
-The event $x^Y = x^{Y'}$ can occur in 4 distinct ways, whose probabilities sum to $P(x^Y = x^{Y'})$.
-At present we work under the assumption that all symbol codes are different and that $Y \neq Y' \implies X^{Y} \neq X^{Y'}.$
+The event $x^Y = x^{Y'}$ can occur in 4 distinct ways, whose probabilities sum to $P(x^Y = x^{Y'})$. At present we assume a negligible chance two sequence or symbol codes are exactly identical.
 
@@
 
-$$= [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t]P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})
+$$= [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t]P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$$
 
@@
 
-\textbf{Sum over all cases}
+\textbf{Sum over cases}
 
@@
 
-When $\gamma = 1$, then $P(x^Y = x^{Y'}) = $
+When $\gamma = 1$ this simplifies to
 
-$$\frac{t+1}{L+1}\frac{t+1}{L'+1}
+$$P(x^Y = x^{Y'}) =  \frac{t+1}{L+1}\frac{t+1}{L'+1}$$
+
+$$
 + \frac{L-t}{L+1}\frac{L'-t}{L'+1}\left[\frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}\right]
@@
 $$+\frac{t+1}{L+1}\frac{L'-t}{L'+1}\left[
-\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}
-\right]
+\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)}+ \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}
+\right]$$
+$$
++\frac{L'-t}{L+1}\frac{t+1}{L'+1}\left[
+\frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} + \left(1 - \frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} \right)\frac{1}{Z}
+\right].
 $$
 
-$$+\frac{L'-t}{L+1}\frac{t+1}{L'+1}\left[
-\frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} + \left(1 - \frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} \right)\frac{1}{Z}
-\right]
-$$
-
-\textbf{OLD STARTS HERE}
-
-This means that when $Y'$ is a starting subsequence of $Y$, i.e. $Y = (y'_1, ..., y'_t, y_{t+1}, ..., y_L)$ then the two sequences' expected similarity is equal to
-
-$$P(x^{Y'} = x^Y) = \frac{1}{(L+1)(t+1)}\left((t+1)^2 + \frac{Z-1}{Z}\mathbf{n}^T_{1:t}\mathbf{n}_{t+1:L} + \frac{(t+1)(L-t)}{Z}\right).$$
-
-\section{Comparison to existing hyperdimensional computing codes}
-
-\bibliographystyle{unsrt}  
-%\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
-%%% and comment out the ``thebibliography'' section.
-
-%%% Comment out this section when you \bibliography{references} is enabled.
-\begin{thebibliography}{1}
-
-\bibitem{kour2014real}
-George Kour and Raid Saabne.
-\newblock Real-time segmentation of on-line handwritten arabic script.
-\newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
-  International Conference on}, pages 417--422. IEEE, 2014.
-
-\bibitem{kour2014fast}
-George Kour and Raid Saabne.
-\newblock Fast classification of handwritten on-line arabic characters.
-\newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
-  International Conference of}, pages 312--318. IEEE, 2014.
-
-\bibitem{hadash2018estimate}
-Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
-  Jacovi.
-\newblock Estimate and replace: A novel approach to integrating deep neural
-  networks with existing applications.
-\newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
-
-\end{thebibliography}
-
-\section{Appendix}
-
-Blah blah
-
 \end{document}

### 2019-09-02 ms/crossing_over.tex
Add seq encoding and decoding algs.

@@
 \usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 
@@
 
-\title{A hyperdimensional sequence code inspired by chromosomal crossover}
+\title{A hyperdimensional code for sequences and trees inspired by chromosomal crossover}
 
@@
 
-\section{Algorithm}
+\section{Problem statement}
 
@@
 
-\subsection{Algorithm}
+\section{Sequence encoding}
 
-Given a dictionary $D$ of $M$ symbols $s^1, ..., s^M$, we will first sample a "C-basis" (crossover basis) $B$ and an auxiliary mask function $\mu$ from $P(B)$ and $P(\mu)$, respectively.
+\textbf{Summary}
+
+Given a dictionary $D$ of $M$ symbols $s^1, ..., s^M$, we first sample a "C-basis" (crossover basis) $B$ and an auxiliary "mask function" $\mu$ from $P(B)$ and $P(\mu)$, respectively.
 Using $B$ and $\mu$, we will then deterministically build codes $\{X^Y\}$ of arbitrary sequences $\{Y = (y_1, ..., y_L)\}$, where $y_t \in D$.
@@
 We reference elements of $\mu$ with the notation $\mu(j, X^A, X^B)$.
+(In practice $\mu$ need not be sampled explicitly, which would take astronomical time; it suffices to use, say, to map $(X^A, X^B)$ using a hash function like SHA-1 to an integer, then sample $N$ times from $\textrm{Uniform}(0, 1)$ from a random number generator seeded with this integer.)
 
-In practice $\mu$ need not be sampled explicitly, which would take astronomical time; it suffices to use, say, to map $(X^A, X^B)$ using a hash function like SHA-1 to an integer, then sample $N$ times from $\textrm{Uniform}(0, 1)$ from a random number generator seeded with this integer.
+\begin{algorithm}
+\caption{Sequence Encoding}
 
-\textbf{Encoding}
+\begin{algorithmic}[0]
+\Function{EncodeSequence}{$Y; B, \mu$}
+\State $X^Y \gets X^*$
+\For{$t \in 1, ..., |Y|$}
 
-...
+\State $X^Y[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma] \gets X^{y_t}[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma]$
+
+\EndFor
+
+\State \textbf{return} $X^Y$
+\EndFunction
+
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-\section{Comparison to existing hyperdimensional computing codes}
+\section{Sequence decoding}
+How to decode a sequence
+
+\begin{algorithm}
+\caption{Sequence Decoding}
+
+\begin{algorithmic}[0]
+\Function{DecodeSequence}{$X^Y; B, \mu$}
+\State $\hat{X}^Y \gets X^*$
+\State $d \gets \textrm{Hamming}(\hat{X}^Y , X^*)$
+\State $t = 1$
+
+\While{$d > 0$}
+
+\State $(i_{best}, \hat{X}^Y_{best}, d_{best}) \gets (Null, Null, 1)$
+
+    \For{$i \in 1, ..., M$}
+
+        \State $\hat{X}^Y_{temp} \gets \hat{X}^Y $
+        \State $\hat{X}^Y_{temp}[\mu(:, \hat{X}^Y, X^i) < 1/(t+1)^\gamma] \gets X^i[\mu(:, \hat{X}^Y, X^i) < 1/(t+1)^\gamma]$
+        \State $d_{temp} \gets \textrm{Hamming}(X^Y, \hat{X}^Y_{temp})$
+
+        \If{$d_{temp} < d_{best}$}
+
+            \State $(i_{best}, \hat{X}^Y_{best}, d_{best}) \gets (i, \hat{X}^Y_{temp}, d_{temp})$
+
+        \EndIf
+    \EndFor
+
+    \If{$d_{best}\geq d$}
+
+        \State \textbf{return} $y_1, ..., y_t$
+
+    \Else
+
+        \State $(y_t, \hat{X}^Y, d) \gets (i_{best}, \hat{X}^Y_{best}, d_{best})$
+        \State $t = t+1$
+
+    \EndIf
+
+\EndWhile
+
+\State \textbf{return} $y_1, ..., y_t$
+\EndFunction
+
+\end{algorithmic}
+\end{algorithm}
+
+\section{Improving decoding with $\gamma$}
+
+\section{Trees}
+
+\begin{algorithm}
+\caption{Tree Encoding}
+
+\begin{algorithmic}[0]
+\Function{EncodeTree}{$T; B, \mu$}
+
+\State $D^T \gets \{i: X^i \textrm{ \textbf{for} } i \in 1, ..., M\}$
+
+\For{$J \in T$}
+
+    \State $X^T \gets D^T[J_1]$
+    \For{$k \in 2, ..., |J|$}
+        \State $mask \gets k/|J| \leq \mu(:, J) \leq (k+1)/|J|$
+        \State $X^T[mask] \gets J_k[mask]$
+    \EndFor
+
+    \State $D^T[J] \gets X^T$
+\EndFor
+
+\State \textbf{return} $D^T[T[end]]$
+\EndFunction
+
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Tree Decoding}
+
+\begin{algorithmic}[0]
+\Function{DecodeTree}{$Y; B, \mu$}
+\State $X^Y \gets X^*$
+\For{$t \in 1, ..., \textrm{len}(Y)$}
+
+\State $X^Y[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma] \gets X^{y_t}[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma]$
+
+\EndFor
+
+\State \textbf{return} $X^Y$
+\EndFunction
+
+\end{algorithmic}
+\end{algorithm}
+
+\section{Mask correlations}
+
+\section{Discussion}
 

### 2019-09-03 ms/crossing_over.tex
Include figs.

@@
 \usepackage[noend]{algpseudocode}
+
+\usepackage{graphicx}
 
@@
 \section{Introduction}
-A core problem in neuroscience and artificial intelligence is how to represent a complex event like a spoken sentence using relatively fixed dimensions.
+A common problem in neuroscience and artificial intelligence is how complex events, like spoken sentences, should be represented in a system of relatively fixed dimension.
 Ideally similar events should have similar representations, and when a new event occurs pre-existing representations shouldn't require substantial rearrangement to prevent confusion.
@@
 
-\State $X^Y[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma] \gets X^{y_t}[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma]$
+\State $mask \gets (\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma)$
+\State $X^Y[mask] \gets X^{y_t}[mask]$
 
@@
 \end{algorithm}
+
+ \begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figs/1_encoding.png}
+    \caption{Sequence encoding algorithm and properties.}
+ \end{figure}
 
@@
 
-    \For{$i \in 1, ..., M$}
+    \For{$i_{temp} \in 1, ..., M$}
 
         \State $\hat{X}^Y_{temp} \gets \hat{X}^Y $
-        \State $\hat{X}^Y_{temp}[\mu(:, \hat{X}^Y, X^i) < 1/(t+1)^\gamma] \gets X^i[\mu(:, \hat{X}^Y, X^i) < 1/(t+1)^\gamma]$
+        \State $ mask \gets (\mu(:, \hat{X}^Y_{temp}, X^{i_{temp}}) < 1/(t+1)^\gamma)$
+        \State $\hat{X}^Y_{temp}[mask] \gets X^{i_{temp}}[mask]$
         \State $d_{temp} \gets \textrm{Hamming}(X^Y, \hat{X}^Y_{temp})$
@@
 
-            \State $(i_{best}, \hat{X}^Y_{best}, d_{best}) \gets (i, \hat{X}^Y_{temp}, d_{temp})$
+            \State $(i_{best}, \hat{X}^Y_{best}, d_{best}) \gets (i_{temp}, \hat{X}^Y_{temp}, d_{temp})$
 
@@
 
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figs/2_decoding.png}
+    \caption{Sequence decoding algorithm and properties.}
+ \end{figure}
+
 \section{Improving decoding with $\gamma$}
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figs/3_gamma.png}
+    \caption{Dependence of sequence codes on $\gamma$.}
+ \end{figure}
 
@@
 
-    \State $X^T \gets D^T[J_1]$
+    \State $X^T \gets D^T[J^1]$
     \For{$k \in 2, ..., |J|$}
-        \State $mask \gets k/|J| \leq \mu(:, J) \leq (k+1)/|J|$
-        \State $X^T[mask] \gets J_k[mask]$
+        \State $mask \gets (k/|J| \leq \mu(:, J) \leq (k+1)/|J|)$
+        \State $X^T[mask] \gets D^T[J^k][mask]$
     \EndFor
@@
 
-\State \textbf{return} $D^T[T[end]]$
+\State \textbf{return} $D^T[T[\textrm{end}]]$
 \EndFunction
@@
 
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figs/4_trees.png}
+    \caption{Tree encoding algorithm and properties.}
+ \end{figure}
+
 \begin{algorithm}
@@
 \begin{algorithmic}[0]
-\Function{DecodeTree}{$Y; B, \mu$}
-\State $X^Y \gets X^*$
-\For{$t \in 1, ..., \textrm{len}(Y)$}
+\Function{DecodeTree}{$X^T; B, \mu$}
+\State $D^T \gets \{i: X^i \textrm{ \textbf{if} } \textrm{Hamming}(X^T, X^i) < d_{exclude} \textrm{ \textbf{for} } i \in 1, ..., M\}$
+\State $T \gets ()$
 
-\State $X^Y[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma] \gets X^{y_t}[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma]$
+\State $d \gets 1$
 
+\While{$d > 0$}
+
+    \State $(J_{best}, \hat{X}^T_{best}, d_{best}) \gets (Null, Null, 1)$
+
+    \For{$J_{temp} \in \{J\}$}
+
+        \State $\hat{X}^T_{temp} \gets D^T[J^1]$
+        \For{$k \in 2, ..., |J_{temp}|$}
+            \State $mask \gets k/|J_{temp}| \leq \mu(:, J_{temp}) \leq (k+1)/|J_{temp}|$
+            \State $\hat{X}^T_{temp}[mask] \gets D^T[J^k_{temp}][mask]$
         \EndFor
 
-\State \textbf{return} $X^Y$
+        \State $d_{temp} \gets \textrm{Hamming}(X^T, \hat{X}^T_{temp})$
+
+        \If{$d_{temp} < d_{best}$}
+            \State $(J_{best}, \hat{X}^T_{best}, d_{best}) \gets (J_{temp}, \hat{X}^T_{temp}, d_{temp})$
+        \EndIf
+    \EndFor
+
+    \If{$d_{best} \geq d$}
+        \State \textbf{return} $T[\textrm{end}]$
+
+    \Else
+        \State $(D^T[J_{best}], d) \gets (\hat{X}^T_{best}, d_{best})$
+        \State $\textrm{Append}(T, J_{best})$
+    \EndIf
+
+\EndWhile
+
+\State \textbf{return} $T[\textrm{end}]$
 \EndFunction
@@
 \section{Mask correlations}
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figs/5_mask_corrs.png}
+    \caption{Robustness to encoding noise through mask correlations.}
+ \end{figure}
 
@@
 
-This follows immediately from \ref{sec:prop-unif-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$.
+This follows from \ref{sec:prop-unif-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$.
 Therefore the statistics of crossover codes do not depend on any aspect about the sequence being encoded.

### 2019-09-05 ms_neurips_2019/ms_neurips_2019.tex
Start ms for neurips 2019 workshop.

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
     \usepackage[final]{neurips_2019}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\usepackage[pdftex]{graphicx}
% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{A crossover code for high-dimensional composition}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Rich Pang\\
  Computational Neuroscience Center\\
  University of Washington\\
  Seattle, WA 98195\\
  \texttt{rpang at uw dot edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
We present a novel method for encoding compositional information by recursively interweaving high-dimensional (HD) random vectors. Inspired by chromosomal crossover, a fraction of one vector's components are masked out and replaced by those from another, via a context-dependent mask. Unlike most HD computing schemes, "crossover" codes highly overlap their base elements' and sub-structures' codes yet without sacrificing relational information, allowing direct element recovery and decoding by greedy reconstruction. Crossover is mathematically tractable and has several properties desirable for robust, flexible representation.
\end{abstract}

\section{Introduction}

A common problem faced by intelligent systems is how to encode objects of variable complexity in fixed dimensions. Ideally, similar objects should have similar codes, new object codes should not degrade existing ones, and there should be no hard object complexity limit. Distributing codes across high-dimensional (HD) vectors yields robustness and potential flexibility and generalization \cite{Hinton:1984, Mikolov:2013}, but how does one encode order or binding relations (key to composition) in such a code, e.g. to distinguish AB vs BA or (AB, CD) vs (AC, CB), respectively? While trained networks have some implicit capacity for this \cite{language network}, it is unclear (1) how such networks might encode complexity beyond that in their training data, and (2) how to quantify internal network representations. An alternative approach, known as HD computing (HDC), is to assign \textit{random} vector codes to a set "base" elements of the data (e.g. symbols), and rules for composing these into richer data structures (e.g. words) \cite{Plate, Kanerva, Gayler}. HDC's advantages are (1) random HD vectors are nearly orthogonal with high probability, enabling a large element codebook, (2) there is no hard limit on data complexity, and (3) encoding is often mathematically tractable. Thus, while a useful system would of course require some training on environment statistics, HDC provides a compelling blueprint for representational flexibility.

Most HDC schemes are based on relational codes that little overlap their element codes (e.g. $AB$'s code is distinct from $A$'s and $B$'s), yet from which the relations can be readily recovered. Examples of typical relation-encoding operations include circular convolution \cite{Plate}, permutation \cite{Sahlgren, Gayler}, or matrix multiplication \cite{Gosmann}. Because of this non-overlap, elements cannot be directly recovered from the relation code (although see \cite{Rachkovskij:2001}), and such a scheme conflicts with our desire that $AB$ should resemble $A$ more than $C$. Here, we present a new way (to our knowledge) to encode relations (and arbitrary compositions), where two vectors $X^A$ and $X^B$ are ordered or bound by masking out a fraction of $X^A$'s components and replacing them with those from $X^B$, analogous to chromosomal crossover, with relational information encoded in a context-dependent mask $\mu(X^A, X^B)$. "Crossover" composition codes highly overlap with the substructures and base elements used to build them, allowing direct recovery of elements and decoding by greedy reconstruction. Further, they distribute information evenly across components and have statistics invariant to object complexity, suggesting utility for robust intelligent systems. We focus on sequences first, then generalize to binding and trees.

\section{Sequence encoding}

Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states, $x_j \in G \equiv \{1,...,Z\}$, and let $d(X^A, X^B)$ be the Hamming distance from $X^A$ to $X^B$. For a symbol sequence $Y = (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, we establish a mapping from $Y$ to $X^Y$ as follows:

First sample a codebook $C \equiv \{X^1, ..., X^M, X^*\}$, where $X^*$ is a "start code", and $x^j$ are i.i.d. across all components and codes and sampled uniformly from $G$. Next sample a mask function $\mu$, an $N \times M^{2N}$ matrix with i.i.d. real-valued elements from $\textrm{Uniform}(0, 1)$, which we reference as $\mu_j(X^A, X^B)$. In practice, columns of $\mu$ can be sampled "as needed" with a hash function.

Using $C$ and $\mu$ we construct sequence codes as described in Algorithm 1 and Figure 1:, symbol codes are recursively woven into the constructed code by replacing a fraction of the current code's components with those from the symbol code; which specific components are replaced depends on the current and symbol codes (the context). We decrease the fraction as $1/(t+1)^\gamma$, for $\gamma \geq 1$ to prevent exponential decay of the first symbol's similarity to the final code.

\begin{algorithm}
\caption{Sequence Encoding}

\begin{algorithmic}[0]
\Function{CrossoverSeq}{$X^A, X^B; t$}

\State $mask \gets (\mu(X^A, X^B) < 1/(t+1)^\gamma) \quad$  \# context-dependent mask w/ about $1/(t+1)^\gamma$ 1's
\State $X^{AB} \gets X^A$
\State $X^{AB}[mask] \gets X^B[mask] \quad$  \# replace $X^A$'s masked components $X^B$'s

\State \textbf{return} $X^{AB}$
\EndFunction

\Function{EncodeSequence}{$Y$}
\State $X^Y \gets X^*$
\For{$t \in 1, ..., |Y|$}

\State $X^Y \gets CrossoverSeq(X^Y, X^{y_t}, t)$

\EndFor

\State \textbf{return} $X^Y$
\EndFunction

\end{algorithmic}
\end{algorithm}

\begin{figure}
  \centering
  \includegraphics{1_sequences.pdf}
  \caption{Sequence encoding via crossover ($N=1024, Z=8, M=26, \gamma=1$). (a) Example encoding of CAFE; colors are states, inscribed symbols are visual aids, not used by the algorithm. (b) Sequence-symbol code distances. (c) State distributions for a short and long sequence. (d, e) Distances of CAFE's and CICADA's code to other sequence codes. (f) Decoding vs code corruption.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{2_gamma.pdf}
  \caption{Crossover modifications. (a) Distance of *REFEREE to symbols/sequences for two $\gamma$. (b) Decoding vs. $\gamma$ for several sequences. (c) Decoding of *CAFE vs corrupted encoding for i.i.d. and correlated $\mu$. (d) Decoding of *CICADA for i.i.d., correlated $\mu$, and correlated $\mu$ with increased $N$.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{3_trees.pdf}
  \caption{Tree encoding via crossover ($N=1024, Z=8, M=26$). (a) Example encoding of (AB, CD, EF). (b) Similarity between final code and element and sub-tree codes.}
\end{figure}

\section{Discussion}

While an intelligent agent must learn environmental statistics to some extent, it can be desirable to certain computations built-in instead of learned from scratch \cite{Zador:2019}. Flexible composition is particularly useful, since lets an agent compute with novel objects/events if they are composed of familiar elements \cite{Jackendoff, Gayler}: it would benefit an autonomous vehicle to be able to automatically generate and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with any of the constituent elements in isolation. Computationally, basing neural network training protocols on built-in HDC codes for composition \cite{Plate} to learn knowledge graphs yielded more efficient computation, easy training, and high scalability \cite{Nickel:2015}. It stands to reason that improving HDC codes may improve network training further still.

As it was discovered early that sums of HD random vectors could encode \textit{sets}, due to high similarity to their components \cite{Bloom:1970, Anderson:1973, Plate:1994}, most HDC work has focused on relations, like sequences or bindings. Perhaps best known are holographic representations \cite{Plate:1995}, where element vectors are bound by circular convolution. Additional proposed binding operations are base on XOR (for binary vectors) \cite{Kanerva:1996}, permutation \cite{Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}. To encode order, convolution can be slightly modified \cite{Murdock:1983, Plate:1995} or prior to summing each element can first be tagged via permutation \cite{Sahlgren:2008} or binding to location tokens \cite{Kanerva:1996}. While such schemes preserve recoverable relational information, the composite vectors are dissimilar to their elements, preventing recovery of the elements from their similarity to the composite vector. An exception is "context-dependent thinning" (CDT), where elements are bound by summing them, then "stamping" them with a set of zeros dependent on the sum \cite{Rachkovskij:2001}; while the stamp shape retains binding information, however, information beneath its imprint is discarded, partially degrading the desired similarities; and it is not clear how CDT could be used to construct recursive sequence codes without early elements' similarities decaying exponentially. The scheme presented in this paper, termed "crossover", overcomes these limitations, preserving similarity to element and sub-structure codes without sacrificing relational information; further, the sequence code automatically prevents early element similarities from being lost, improving decodability. Beyond this, crossover codes have i.i.d. information across the vector components, and code statistics are invariant to object complexity. Crossover might thus be a useful blueprint for a robust cognitive system capable of working with arbitrary objects.

We close by discussing HDC's potential implementation in biological neural networks. Of key interest is the working memory (WM) system, which must generate and maintain representations of novel objects (e.g. a never-before-heard question) on the fly. It has been hypothesized that information persists in WM through persistent patterns of neural activity \cite{Chaudhuri Nat Neuro}. While certain networks can support the combinatorial persistent patterns required for composition \cite{papers from CSHL application}, how they might carry meaning is little understood. HDC suggests an answer: by decomposing such patterns into independent components and assigning these to the components of an HDC vector, one could map the space of HDC representations to plausible neural activity, so that the latter inherited HDC's representational flexibility. This could provide a robust, expressive reservoir of neural representations that could be manipulated by recurrent computations.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All acknowledgments
go at the end of the paper. Do not include acknowledgments in the anonymized
submission, only in the final paper.

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references. {\bf Remember that you can use more than eight
  pages as long as the additional pages contain \emph{only} cited references.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2019-09-05 ms_neurips_2019/ms_neurips_2019.tex
Finish first draft.

@@
 \begin{abstract}
-We present a novel method for encoding compositional information by recursively interweaving high-dimensional (HD) random vectors.
+We present a novel method for encoding compositional information by recursively weaving together high-dimensional (HD) random vectors.
 Inspired by chromosomal crossover, a fraction of one vector's components are masked out and replaced by those from another, via a context-dependent mask.
-Unlike most HD computing schemes, "crossover" codes highly overlap their base elements' and sub-structures' codes yet without sacrificing relational information, allowing direct element recovery and decoding by greedy reconstruction.
+Unlike most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes yet without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
 Crossover is mathematically tractable and has several properties desirable for robust, flexible representation.
@@
 A common problem faced by intelligent systems is how to encode objects of variable complexity in fixed dimensions.
-Ideally, similar objects should have similar codes, new object codes should not degrade existing ones, and there should be no hard object complexity limit.
-Distributing codes across high-dimensional (HD) vectors yields robustness and potential flexibility and generalization \cite{Hinton:1984, Mikolov:2013}, but how does one encode order or binding relations (key to composition) in such a code, e.g.
+Ideally, similar objects should have similar codes, new objects should not require rearranging existing codes, and there should be no limit on object complexity.
+Distributing codes across high-dimensional (HD) vectors yields robustness and potential flexibility and generalization \cite{Hinton:1984, Mikolov:2013}, but how does one impose compositional relations, like \textit{order} or \textit{binding} in such a code, e.g.
 to distinguish AB vs BA or (AB, CD) vs (AC, CB), respectively?
-While trained networks have some implicit capacity for this \cite{language network}, it is unclear (1) how such networks might encode complexity beyond that in their training data, and (2) how to quantify internal network representations.
-An alternative approach, known as HD computing (HDC), is to assign \textit{random} vector codes to a set "base" elements of the data (e.g.
+While trained networks have some implicit capacity for this \cite{Bahdanau:2014, Luong:2015, Wu:2016}, it is unclear (1) how such networks might encode complexity beyond that of their training data, and (2) how to quantify internal network function \cite{Lipton:2017}.
+An alternative approach, known as HD computing (HDC), is to assign \textit{random} vector codes to a set of "base" elements of the data (e.g.
 symbols), and rules for composing these into richer data structures (e.g.
-words) \cite{Plate, Kanerva, Gayler}.
-HDC's advantages are (1) random HD vectors are nearly orthogonal with high probability, enabling a large element codebook, (2) there is no hard limit on data complexity, and (3) encoding is often mathematically tractable.
-Thus, while a useful system would of course require some training on environment statistics, HDC provides a compelling blueprint for representational flexibility.
+words) \cite{Plate:1995, Kanerva:2009, Gayler}.
+HDC's advantages are (1) random HD vectors have low overlap with high probability, enabling a large element codebook, (2) there is no hard limit on data complexity, and (3) encoding is often mathematically tractable.
+While a useful system would of course require some training on environment statistics, HDC provides a compelling blueprint for representational flexibility.
 
-Most HDC schemes are based on relational codes that little overlap their element codes (e.g.
-$AB$'s code is distinct from $A$'s and $B$'s), yet from which the relations can be readily recovered.
-Examples of typical relation-encoding operations include circular convolution \cite{Plate}, permutation \cite{Sahlgren, Gayler}, or matrix multiplication \cite{Gosmann}.
-Because of this non-overlap, elements cannot be directly recovered from the relation code (although see \cite{Rachkovskij:2001}), and such a scheme conflicts with our desire that $AB$ should resemble $A$ more than $C$.
-Here, we present a new way (to our knowledge) to encode relations (and arbitrary compositions), where two vectors $X^A$ and $X^B$ are ordered or bound by masking out a fraction of $X^A$'s components and replacing them with those from $X^B$, analogous to chromosomal crossover, with relational information encoded in a context-dependent mask $\mu(X^A, X^B)$.
-"Crossover" composition codes highly overlap with the substructures and base elements used to build them, allowing direct recovery of elements and decoding by greedy reconstruction.
-Further, they distribute information evenly across components and have statistics invariant to object complexity, suggesting utility for robust intelligent systems.
-We focus on sequences first, then generalize to binding and trees.
+Most HDC schemes are based on codes in which relations among elements can be readily recovered, but only if one provides certain elements themselves as cues.
+Typical relation-encoding operations include circular convolution \cite{Plate}, permutation \cite{Sahlgren, Gayler}, or matrix multiplication \cite{Gosmann}, which preserve relational information by creating relation codes that little overlap their elements (e.g.
+AB's code is distinct from A's and B's).
+This low overlap means elements cannot be directly recovered from the relation code (although see \cite{Rachkovskij:2001}), and it conflicts with our desire that AB should resemble A and B more than C.
+Here, we present a new way (to our knowledge) to encode compositional relations, where two HD vectors $X$ and $X'$ are ordered or bound by masking out a fraction of $X$'s components and replacing them with those from $X'$, analogous to chromosomal crossover, and with relational information encoded in the context-dependent mask $\mu(X, X')$.
+As a result, "crossover" composition codes highly overlap with the substructures and elements used to build them, allowing direct recovery of base elements and decoding by greedy reconstruction.
+Further, they distribute information evenly across components and have marginal statistics invariant to object complexity, suggesting utility for robust intelligent systems.
+We describe crossover for sequences first, then generalize to binding and trees.
 
-\section{Sequence encoding}
+\section{Results}
 
-Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states, $x_j \in G \equiv \{1,...,Z\}$, and let $d(X^A, X^B)$ be the Hamming distance from $X^A$ to $X^B$.
+Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states, $x_j \in G \equiv \{1,...,Z\}$, and let $d(X, X')$ be the Hamming distance from $X$ to $X'$.
 For a symbol sequence $Y = (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, we establish a mapping from $Y$ to $X^Y$ as follows:
@@
 Next sample a mask function $\mu$, an $N \times M^{2N}$ matrix with i.i.d.
-real-valued elements from $\textrm{Uniform}(0, 1)$, which we reference as $\mu_j(X^A, X^B)$.
+real-valued elements from $\textrm{Uniform}(0, 1)$, which we reference as $\mu_j(X, X')$.
 In practice, columns of $\mu$ can be sampled "as needed" with a hash function.
 
-Using $C$ and $\mu$ we construct sequence codes as described in Algorithm 1 and Figure 1:, symbol codes are recursively woven into the constructed code by replacing a fraction of the current code's components with those from the symbol code; which specific components are replaced depends on the current and symbol codes (the context).
-We decrease the fraction as $1/(t+1)^\gamma$, for $\gamma \geq 1$ to prevent exponential decay of the first symbol's similarity to the final code.
+Using $C$ and $\mu$ we encode sequences as in Alg 1 and Fig 1a: symbol codes are recursively woven into the final code by replacing a fraction of the current code's components with those from the symbol code; which components are replaced encodes sequence order.
+We decrease the fraction as $1/(t+1)^\gamma$, for $\gamma \geq 1$ to prevent exponential decay of the first symbol's presence in the final code.
 
@@
 \caption{Sequence Encoding}
+\label{alg:1}
 
 \begin{algorithmic}[0]
-\Function{CrossoverSeq}{$X^A, X^B; t$}
+\Function{CrossoverSeq}{$X, X'; t$} $\rightarrow X''$
 
-\State $mask \gets (\mu(X^A, X^B) < 1/(t+1)^\gamma) \quad$  \# context-dependent mask w/ about $1/(t+1)^\gamma$ 1's
-\State $X^{AB} \gets X^A$
-\State $X^{AB}[mask] \gets X^B[mask] \quad$  \# replace $X^A$'s masked components $X^B$'s
+\State $mask \gets (\mu(X, X') < 1/(t+1)^\gamma) \quad$  \# context-dependent mask with about $1/(t+1)^\gamma$ 1's
+\State $X'' \gets X$
+\State $X''[mask] \gets X'[mask] \quad$  \# replace $X$'s masked components with those from $X'$
 
-\State \textbf{return} $X^{AB}$
 \EndFunction
 
-\Function{EncodeSequence}{$Y$}
+\State
+
+\Function{EncodeSequence}{$Y$} $\rightarrow X^Y$
+
 \State $X^Y \gets X^*$
@@
 
-\State $X^Y \gets CrossoverSeq(X^Y, X^{y_t}, t)$
+\State $X^Y \gets \textrm{CrossoverSeq}(X^Y, X^{y_t}, t)$
 
@@
 
-\State \textbf{return} $X^Y$
 \EndFunction
@@
 (d, e) Distances of CAFE's and CICADA's code to other sequence codes.
-(f) Decoding vs code corruption.}
+(f) Decoding vs code corruption.
+All plots here and elsewhere were made via 100 samples of $C, \mu$.
+"*" prefixes indicate sequences.}
+  \label{fig:1}
 \end{figure}
 
+For $\gamma = 1$, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$; $n^i$ is how often $i$ appears in $Y$.
+I.e.
+for large $Z$, mean symbol-sequence similarity (over $C, \mu$) scales with the symbol's count within the sequence (Fig 1b), with variance $\sim 1/N$.
+As exemplified by Fig 1c, crossover code components are uniform over $G$, i.i.d.
+for a given $Y$, and invariant to sequence statistics (e.g.
+length).
+The proof follows from the i.i.d.
+nature of $C, \mu$ and the interchangeability of component states, and suggests crossover as a robust, scale-invariant encoding.
+
+For two sequences $Y$ and $Y'$ sharing their first $t$ symbols, $d(X^Y, X^{Y'})$ is also a scaled Binomial variable, and $q$ can be computed exactly for arbitrary $\gamma$; we present the $\gamma = 1$ case here.
+Let $\mathbf{n}_u$ be the length-$M$ symbol count vector for $Y_{1:t} = Y'_{1:t}$ with $\mathbf{1}^T\mathbf{n}_u = t$, and $\mathbf{n}_v$ and $\mathbf{n}_{v'}$ be the symbol count vectors for $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively, with $\mathbf{1}^T\mathbf{n}_{v} = L - t$ and $\mathbf{1}^T\mathbf{n}_{v'} = L' - t$.
+Then:
+$$q \equiv P(x_j^Y = x_j^{Y'}) = \frac{t+1}{L+1}\frac{t+1}{L'+1} + \frac{1}{L+1}\frac{1}{L'+1}\left[\mathbf{n}^T_v\mathbf{n}_{v'} + \left((L-t)(L'-t) - \mathbf{n}^T_v\mathbf{n}_{v'} \right)\frac{1}{Z}\right]$$
+$$+\frac{1}{L+1}\frac{1}{L'+1}\left[
+\mathbf{n}^T_{u}\mathbf{n}_{v'} + \left((t+1)(L'-t) - \mathbf{n}^T_{u}\mathbf{n}_{v'} \right)\frac{1}{Z}
++ \mathbf{n}^T_{v}\mathbf{n}_{u} + \left((L-t)(t+1) - \mathbf{n}^T_{v}\mathbf{n}_{u} \right)\frac{1}{Z}
+\right]$$
+Consequently, for most $Y$, $X^Y$ overlaps little with $Y$'s anagrams, and more with correct than incorrect starting subsequences (Fig 1d,e): CICADA is closer to CI than CD and to CIC than CIA.
+This is because distinct next-symbol candidates yield different $\mu$, which quickly decreases overlap with $X^Y$ for out-of-order symbols.
+Thus, $Y$ is rapidly decoded by rebuilding $X^Y$ from $X^*$ if each rebuild step maximizes overlap with $X^Y$, until overlap stops increasing.
+Further, given a prior $P(Y)$, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
+across $j$, yielding equivalent robustness to random and targeted attack, so decoding accuracy depends only on the number of corrupted components (Fig 1f).
+
+For $Y$ with many symbol repeats distinct from $y_1$ (e.g.
+REFEREE), $X^*$ crossed over with $X^{y_1}$ can be further from $X^Y$ than $X^*$ crossed with the repeated symbol, impairing decoding.
+Making $\gamma > 1$ fixes this by more heavily weighting early symbols (Fig 2a).
+Surprisingly, although this imbalanced weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared across many sequences (Fig 2b).
+
 \begin{figure}
+\label{fig:2}
   \centering
   \includegraphics{2_gamma.pdf}
-  \caption{Crossover modifications.
+  \caption{Crossover modifications ($N=1024, Z=8, M=26$).
 (a) Distance of *REFEREE to symbols/sequences for two $\gamma$.
@@
 $\gamma$ for several sequences.
-(c) Decoding of *CAFE vs corrupted encoding for i.i.d.
+(c) Decoding of *CAFE vs number of corrupted components at encoding step for i.i.d.
 and correlated $\mu$.
@@
 
+While crossover with i.i.d.
+$\mu$ is robust to noise added to the final code (Fig 1f), it fails when noise is added during encoding (Fig 2c,d), as one altered component yields an entirely different mask.
+This is fixed by introducing correlations into $\mu_j(X, X')$ (here making $\mu$ a function of sparse components of $X, X'$) (Fig 2c,d).
+The tradeoff is decreased decoding accuracy for certain sequences, since correlated masks are less distinct, but this can be corrected via increased $N$.
+
+Beyond sequences, crossover naturally generalizes to nested structures, i.e.
+trees.
+Key to tree encoding is \textit{binding} elements together into groups that can themselves be bound together, yet without erasing the initial group identities, e.g.
+(AB, CD) must not be confused with (AD, CB).
+In crossover, $k$ elements $i_0, ..., i_k$ can be bound through a context-dependent mask function $\mu(X^{i_0}, ....
+X^{i_k})$ (which can be made to commute if order does not matter).
+The output vector then borrows components from $X^{i_0}$ where $0 < \mu < 1/k$, from $X^{i_1}$ where $1/k < \mu < 2/k$, etc.
+These outputs are then bound themselves in the exact same manner (Fig 3a), an operation which can be applied recursively according to the tree depth.
+The result is a vector overlapping more with included than excluded elements, and with included than excluded groups, even if the latter are made from the same base elements (Fig 3b).
+Generally, a tree encoding by crossover will also overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
+Similar to sequence encoding, this means that both base elements and their recursive relationships can be extracted well from the crossover code, and that decoding of the full tree can occur through a recursive, but greedy reconstruction.
+
 \begin{figure}
+\label{fig:3}
   \centering
   \includegraphics{3_trees.pdf}
-  \caption{Tree encoding via crossover ($N=1024, Z=8, M=26$). (a) Example encoding of (AB, CD, EF). (b) Similarity between final code and element and sub-tree codes.}
+  \caption{Tree/binding encoding via crossover ($N=1024, Z=8, M=26$). (a) Example crossover encoding of (AB, CD, EF). (b) Distances of final code to element and bound-pair codes.}
 \end{figure}
-
 \section{Discussion}
 
-While an intelligent agent must learn environmental statistics to some extent, it can be desirable to certain computations built-in instead of learned from scratch \cite{Zador:2019}.
-Flexible composition is particularly useful, since lets an agent compute with novel objects/events if they are composed of familiar elements \cite{Jackendoff, Gayler}: it would benefit an autonomous vehicle to be able to automatically generate and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with any of the constituent elements in isolation.
-Computationally, basing neural network training protocols on built-in HDC codes for composition \cite{Plate} to learn knowledge graphs yielded more efficient computation, easy training, and high scalability \cite{Nickel:2015}.
-It stands to reason that improving HDC codes may improve network training further still.
+While intelligent systems must generally learn the environment statistics to operate, it can be desirable for some features of computation to be built-in automatically \cite{Zador:2019}.
+Flexible composition codes are useful because they provide a substrate for an agent to compute with novel objects/events if they are composed of familiar elements.
+For instance, an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with the isolated scene elements.
+While many trained systems, e.g.
+translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some capacity for compositional encoding (otherwise outputs would be out-of-order), it is unclear how they could handle complexity beyond their training data, and it is notoriously difficult to understand the structure of their internal representations \cite{Zeiler:2013, Montavon:2018}.
+Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
+Basing a training protocol on convolutional binding, for instance, allowed an artificial network to learn knowledge graphs more efficiently and scalably than other methods \cite{Nickel:2015}.
+It stands to reason that identifying robust HDC codes could provide a useful backbone for improving artificial systems, as well as for directing our investigation of neural computation in biology.
 
-As it was discovered early that sums of HD random vectors could encode \textit{sets}, due to high similarity to their components \cite{Bloom:1970, Anderson:1973, Plate:1994}, most HDC work has focused on relations, like sequences or bindings.
-Perhaps best known are holographic representations \cite{Plate:1995}, where element vectors are bound by circular convolution.
-Additional proposed binding operations are base on XOR (for binary vectors) \cite{Kanerva:1996}, permutation \cite{Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}.
-To encode order, convolution can be slightly modified \cite{Murdock:1983, Plate:1995} or prior to summing each element can first be tagged via permutation \cite{Sahlgren:2008} or binding to location tokens \cite{Kanerva:1996}.
-While such schemes preserve recoverable relational information, the composite vectors are dissimilar to their elements, preventing recovery of the elements from their similarity to the composite vector.
-An exception is "context-dependent thinning" (CDT), where elements are bound by summing them, then "stamping" them with a set of zeros dependent on the sum \cite{Rachkovskij:2001}; while the stamp shape retains binding information, however, information beneath its imprint is discarded, partially degrading the desired similarities; and it is not clear how CDT could be used to construct recursive sequence codes without early elements' similarities decaying exponentially.
-The scheme presented in this paper, termed "crossover", overcomes these limitations, preserving similarity to element and sub-structure codes without sacrificing relational information; further, the sequence code automatically prevents early element similarities from being lost, improving decodability.
-Beyond this, crossover codes have i.i.d.
-information across the vector components, and code statistics are invariant to object complexity.
-Crossover might thus be a useful blueprint for a robust cognitive system capable of working with arbitrary objects.
+"Crossover" shares advantages with existing HDC schemes but is distinct in key ways also.
+Like other schemes, crossover preserves relational information among elements, which is readily recoverable from the final code \cite{Kanerva:2009}.
+Further, relation codes can be reused as elements themselves, allowing arbitrary deep recursive composition \cite{Kanerva:2009}; in crossover one can add CAFE to the dictionary (perhaps with minor noise added), enabling encoding of longer sequences made from shorter ones.
+To encode relations, however, previous schemes have relied on circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1996}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yields relation codes that little overlap their elements.
+While these distinguish different compositions of the same elements and allow cue-based recall, base elements cannot be identified directly from their overlap with the relation code, and instead require additional storage.
+An exception is "context-dependent thinning" (CDT), which stores relations via the union of the elements, then "stamping" the code with a context-dependent set of zeros \cite{Rachkovskij:2001}; while the stamp shape retains relational information, however, information beneath its imprint is discarded, and it is not clear how CDT could construct recursive sequence codes without early elements decaying exponentially.
+Crossover overcomes these limitations, as both elements and their relations can be read out directly, and in sequences no element overlaps decay exponentially.
+Further, crossover codes have i.i.d.
+information across code components, and code statistics are invariant to object complexity, suggesting they might be a useful representation for a robust intelligent system.
 
-We close by discussing HDC's potential implementation in biological neural networks.
-Of key interest is the working memory (WM) system, which must generate and maintain representations of novel objects (e.g.
-a never-before-heard question) on the fly.
-It has been hypothesized that information persists in WM through persistent patterns of neural activity \cite{Chaudhuri Nat Neuro}.
-While certain networks can support the combinatorial persistent patterns required for composition \cite{papers from CSHL application}, how they might carry meaning is little understood.
-HDC suggests an answer: by decomposing such patterns into independent components and assigning these to the components of an HDC vector, one could map the space of HDC representations to plausible neural activity, so that the latter inherited HDC's representational flexibility.
-This could provide a robust, expressive reservoir of neural representations that could be manipulated by recurrent computations.
+Crossover is also conceptually distinct from its predecessors since one need not define arithmetical operations over code components.
+Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Kanerva:1996, Gayler:1998, Rachkovskij:2001, Sahlgren:2008, Gosmann:2019}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
+Crossover thus outlines a distributed computation substrate that can exist outside vector spaces, and instead within more generic metric spaces.
+This is relevant to biological implementation, as it allows one to envision code components not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank.
+For instance, each component of a crossover code could correspond to the attractor state of a 100-neuron Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree.
+Such could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code overlap statistics to be usefully read out and manipulated by downstream and recurrent computations.
 

### 2019-09-08 ms_neurips_2019/ms_neurips_2019.tex
Fix refs and start editing.

@@
 \usepackage[noend]{algpseudocode}
+% \usepackage[superscript,biblabel]{cite}
 
@@
 Inspired by chromosomal crossover, a fraction of one vector's components are masked out and replaced by those from another, via a context-dependent mask.
-Unlike most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes yet without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
+Contrasting most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes yet without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
 Crossover is mathematically tractable and has several properties desirable for robust, flexible representation.
@@
 Ideally, similar objects should have similar codes, new objects should not require rearranging existing codes, and there should be no limit on object complexity.
-Distributing codes across high-dimensional (HD) vectors yields robustness and potential flexibility and generalization \cite{Hinton:1984, Mikolov:2013}, but how does one impose compositional relations, like \textit{order} or \textit{binding} in such a code, e.g.
-to distinguish AB vs BA or (AB, CD) vs (AC, CB), respectively?
-While trained networks have some implicit capacity for this \cite{Bahdanau:2014, Luong:2015, Wu:2016}, it is unclear (1) how such networks might encode complexity beyond that of their training data, and (2) how to quantify internal network function \cite{Lipton:2017}.
+Distributing codes across high-dimensional (HD) vectors yields robustness and potential flexibility and generalization \cite{Hinton:1984, Mikolov:2013}, but how does one store compositional relations, like order or binding in such a code, e.g.
+to distinguish AB vs BA or ((A,B), (C,D)) vs ((A,D), (C,B)), respectively?
+While trained networks have some implicit capacity for this \cite{Bahdanau:2014, Luong:2015, Wu:2016}, it is unclear (1) how such networks might encode complexity beyond that of their training data, and (2) how to quantify internal network representations \cite{Lipton:2016}.
 An alternative approach, known as HD computing (HDC), is to assign \textit{random} vector codes to a set of "base" elements of the data (e.g.
 symbols), and rules for composing these into richer data structures (e.g.
-words) \cite{Plate:1995, Kanerva:2009, Gayler}.
-HDC's advantages are (1) random HD vectors have low overlap with high probability, enabling a large element codebook, (2) there is no hard limit on data complexity, and (3) encoding is often mathematically tractable.
-While a useful system would of course require some training on environment statistics, HDC provides a compelling blueprint for representational flexibility.
+words) \cite{Plate:1995, Kanerva:2009, Gayler:2004}.
+HDC's advantages are (1) random HD vectors have low overlap with high probability, enabling large element codebooks, (2) there is no hard limit on data complexity, and (3) encoding is often analytically tractable.
+While most systems would of course require some training on environment statistics, HDC provides a compelling blueprint for representational flexibility.
 
-Most HDC schemes are based on codes in which relations among elements can be readily recovered, but only if one provides certain elements themselves as cues.
-Typical relation-encoding operations include circular convolution \cite{Plate}, permutation \cite{Sahlgren, Gayler}, or matrix multiplication \cite{Gosmann}, which preserve relational information by creating relation codes that little overlap their elements (e.g.
+Most HDC schemes use codes in which relations among elements can be easily recovered, but only if one provides certain elements as cues (e.g.
+cue A recovers B from ((A,B),(C,D))).
+Typical relation-encoding operations include circular convolution \cite{Plate:1995}, permutation \cite{Gayler:1998, Sahlgren:2008}, or matrix multiplication \cite{Gosmann:2019}, which create relational codes that little overlap their elements' codes (e.g.
 AB's code is distinct from A's and B's).
@@
 (b) Sequence-symbol code distances.
-(c) State distributions for a short and long sequence.
+(c) State distributions for example short and long sequence for one $C, \mu$ sample.
 (d, e) Distances of CAFE's and CICADA's code to other sequence codes.
 (f) Decoding vs code corruption.
+"x"s indicate theoretical predictions and bars average simulation results.
 All plots here and elsewhere were made via 100 samples of $C, \mu$.
@@
 \right]$$
-Consequently, for most $Y$, $X^Y$ overlaps little with $Y$'s anagrams, and more with correct than incorrect starting subsequences (Fig 1d,e): CICADA is closer to CI than CD and to CIC than CIA.
-This is because distinct next-symbol candidates yield different $\mu$, which quickly decreases overlap with $X^Y$ for out-of-order symbols.
-Thus, $Y$ is rapidly decoded by rebuilding $X^Y$ from $X^*$ if each rebuild step maximizes overlap with $X^Y$, until overlap stops increasing.
-Further, given a prior $P(Y)$, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
+For most $Y$, $X^Y$ overlaps little with $Y$'s anagrams, and more with correct than incorrect starting subsequences, e.g.
+codes for CI and CIC are closer to CICADA's code than codes for CD and CIA, respectively (Fig 1d,e).
+This is because different next-symbol candidates yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols.
+Consequently, $Y$ can greedily decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step the symbol maximizing overlap with $X^Y$ when woven into the current code, then weaving it in.
+Of note, decoding as such is noise-robust, and when multiple $Y$ are given via a prior $P(Y)$, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
 across $j$, yielding equivalent robustness to random and targeted attack, so decoding accuracy depends only on the number of corrupted components (Fig 1f).
@@
 REFEREE), $X^*$ crossed over with $X^{y_1}$ can be further from $X^Y$ than $X^*$ crossed with the repeated symbol, impairing decoding.
-Making $\gamma > 1$ fixes this by more heavily weighting early symbols (Fig 2a).
-Surprisingly, although this imbalanced weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared across many sequences (Fig 2b).
+Making $\gamma > 1$, however, fixes this by more heavily weighting early symbols (Fig 2a).
+Surprisingly, although this front-loaded weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared across many sequences (Fig 2b).
+In general, decoding accuracy decreases with L and increases with N, as the variance of $P(x_j^Y = x_j^Y')$, with $Y'$ a starting subsequence of $Y$ scales as $1/N$.
+Notably, decoding accuracy depends little of $M$, the dictionary size, due to the large expected distances between HD random vectors.
 
@@
 
-Beyond sequences, crossover naturally generalizes to nested structures, i.e.
-trees.
-Key to tree encoding is \textit{binding} elements together into groups that can themselves be bound together, yet without erasing the initial group identities, e.g.
-(AB, CD) must not be confused with (AD, CB).
-In crossover, $k$ elements $i_0, ..., i_k$ can be bound through a context-dependent mask function $\mu(X^{i_0}, ....
-X^{i_k})$ (which can be made to commute if order does not matter).
-The output vector then borrows components from $X^{i_0}$ where $0 < \mu < 1/k$, from $X^{i_1}$ where $1/k < \mu < 2/k$, etc.
-These outputs are then bound themselves in the exact same manner (Fig 3a), an operation which can be applied recursively according to the tree depth.
-The result is a vector overlapping more with included than excluded elements, and with included than excluded groups, even if the latter are made from the same base elements (Fig 3b).
-Generally, a tree encoding by crossover will also overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
-Similar to sequence encoding, this means that both base elements and their recursive relationships can be extracted well from the crossover code, and that decoding of the full tree can occur through a recursive, but greedy reconstruction.
+Beyond sequences, crossover naturally generalizes to nested structures, or trees.
+Key to encoding trees is \textit{binding} elements together into groups that can themselves be bound, yet without forgetting initial group identities, e.g.
+((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)).
+In crossover, $k$ elements $i_0, ..., i_k$ can be bound into a group by using a context-dependent mask function $\mu(X^{i_0}, ....
+X^{i_k})$ (which can commute if within-group order does not matter).
+The vector for the group takes components from $X^{i_0}$ where $0 < \mu < 1/k$, from $X^{i_1}$ where $1/k < \mu < 2/k$, etc.
+This vector can then be bound itself in the exact same manner (Fig 3a), allowing one to recursively encode a tree of arbitrary depth.
+The result is a vector overlapping more with elements within vs not in the tree, and with groups within vs not in tree, even if the latter contain same base elements (Fig 3b).
+Generally, a tree encoding by crossover will tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
+Similar to sequence encoding, this means that both base elements and their recursive relationships can be extracted via their overlap with the crossover code, and that decoding of the full tree can occur through a recursive, but greedy reconstruction.
 
@@
   \includegraphics{3_trees.pdf}
-  \caption{Tree/binding encoding via crossover ($N=1024, Z=8, M=26$). (a) Example crossover encoding of (AB, CD, EF). (b) Distances of final code to element and bound-pair codes.}
+  \caption{Tree/binding encoding via crossover ($N=1024, Z=8, M=26$). (a) Example crossover encoding of ((A,B), (C,D), (E,F)). (b) Distances of final code to element and bound-pair codes.}
 \end{figure}
@@
 
-While intelligent systems must generally learn the environment statistics to operate, it can be desirable for some features of computation to be built-in automatically \cite{Zador:2019}.
-Flexible composition codes are useful because they provide a substrate for an agent to compute with novel objects/events if they are composed of familiar elements.
-For instance, an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with the isolated scene elements.
+While intelligent systems must learn environment statistics to operate successfully, it can useful for certain computational features to be built in \cite{Zador:2019}.
+Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects/events if they are made from familiar elements: an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only learned the isolated scene elements.
 While many trained systems, e.g.
-translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some capacity for compositional encoding (otherwise outputs would be out-of-order), it is unclear how they could handle complexity beyond their training data, and it is notoriously difficult to understand the structure of their internal representations \cite{Zeiler:2013, Montavon:2018}.
+translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}).
 Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
-Basing a training protocol on convolutional binding, for instance, allowed an artificial network to learn knowledge graphs more efficiently and scalably than other methods \cite{Nickel:2015}.
-It stands to reason that identifying robust HDC codes could provide a useful backbone for improving artificial systems, as well as for directing our investigation of neural computation in biology.
+Basing training on convolutional binding, for instance, allowed an artificial network to learn knowledge graphs more efficiently and scalably than other methods \cite{Nickel:2016}.
+It stands to reason that identifying robust HDC codes could provide a useful backbone to improve artificial systems, as well as for directing our investigation of neural computation in biology.
 
@@
 Further, relation codes can be reused as elements themselves, allowing arbitrary deep recursive composition \cite{Kanerva:2009}; in crossover one can add CAFE to the dictionary (perhaps with minor noise added), enabling encoding of longer sequences made from shorter ones.
-To encode relations, however, previous schemes have relied on circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1996}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yields relation codes that little overlap their elements.
+To encode relations, however, previous schemes have relied on circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1994}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yields relation codes that little overlap their elements.
 While these distinguish different compositions of the same elements and allow cue-based recall, base elements cannot be identified directly from their overlap with the relation code, and instead require additional storage.
@@
 Crossover overcomes these limitations, as both elements and their relations can be read out directly, and in sequences no element overlaps decay exponentially.
-Further, crossover codes have i.i.d.
-information across code components, and code statistics are invariant to object complexity, suggesting they might be a useful representation for a robust intelligent system.
+Further, crossover yields i.i.d.
+information across code components and code statistics invariant to object complexity, suggesting it may be a useful representation for a robust intelligent system.
 
 Crossover is also conceptually distinct from its predecessors since one need not define arithmetical operations over code components.
-Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Kanerva:1996, Gayler:1998, Rachkovskij:2001, Sahlgren:2008, Gosmann:2019}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
+Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Kanerva:1994, Gayler:1998, Rachkovskij:2001, Sahlgren:2008, Gosmann:2019}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
 Crossover thus outlines a distributed computation substrate that can exist outside vector spaces, and instead within more generic metric spaces.
-This is relevant to biological implementation, as it allows one to envision code components not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank.
+This is relevant to biological implementation, as it allows one to envision $x_j$ not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank.
 For instance, each component of a crossover code could correspond to the attractor state of a 100-neuron Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree.
@@
 
-\section*{References}
+\bibliographystyle{unsrt}  
+\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
 
-References follow the acknowledgments. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references. {\bf Remember that you can use more than eight
-  pages as long as the additional pages contain \emph{only} cited references.}
-\medskip
+%%% and comment out the ``thebibliography'' section.
 
-\small
+% \section*{References}
 
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
+% References follow the acknowledgments. Use unnumbered first-level heading for
+% the references. Any choice of citation style is acceptable as long as you are
+% consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
+% when listing the references. {\bf Remember that you can use more than eight
+%   pages as long as the additional pages contain \emph{only} cited references.}
+% \medskip
 
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
+% \small
 
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
+% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
+% connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
+% (eds.), {\it Advances in Neural Information Processing Systems 7},
+% pp.\ 609--616. Cambridge, MA: MIT Press.
+
+% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
+%   Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
+% TELOS/Springer--Verlag.
+
+% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
+% recall at excitatory recurrent synapses and cholinergic modulation in rat
+% hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
 

### 2019-09-13 ms_neurips_2019/ms_neurips_2019.tex
Go through next revision.

@@
 \begin{abstract}
-We present a novel method for encoding compositional information by recursively weaving together high-dimensional (HD) random vectors.
-Inspired by chromosomal crossover, a fraction of one vector's components are masked out and replaced by those from another, via a context-dependent mask.
-Contrasting most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes yet without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
+We present a novel way to encode compositional information in high-dimensional (HD) vectors.
+Inspired by chromosomal crossover, random HD vectors are recursively interwoven, with a fraction of one vector's components masked out and replaced by those from another using a context-dependent mask.
+Unlike most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes, allowing fast element readout without sacrificing relational information, and decoding by greedy reconstruction.
 Crossover is mathematically tractable and has several properties desirable for robust, flexible representation.
@@
 Ideally, similar objects should have similar codes, new objects should not require rearranging existing codes, and there should be no limit on object complexity.
-Distributing codes across high-dimensional (HD) vectors yields robustness and potential flexibility and generalization \cite{Hinton:1984, Mikolov:2013}, but how does one store compositional relations, like order or binding in such a code, e.g.
+Distributing codes across high-dimensional (HD) vectors yields potential robustness, flexibility, and generalization \cite{Hinton:1984, Mikolov:2013}; but how does one store compositional relations, like order or binding in such a code, e.g.
 to distinguish AB vs BA or ((A,B), (C,D)) vs ((A,D), (C,B)), respectively?
-While trained networks have some implicit capacity for this \cite{Bahdanau:2014, Luong:2015, Wu:2016}, it is unclear (1) how such networks might encode complexity beyond that of their training data, and (2) how to quantify internal network representations \cite{Lipton:2016}.
-An alternative approach, known as HD computing (HDC), is to assign \textit{random} vector codes to a set of "base" elements of the data (e.g.
-symbols), and rules for composing these into richer data structures (e.g.
+While trained systems like deep networks have some implicit capacity for this \cite{Bahdanau:2014, Luong:2015, Wu:2016}, it is unclear how these might handle complexity beyond that of their training data and challenging to quantify internal distributed representations \cite{Lipton:2016}.
+An alternative approach, known as HD computing (HDC), is to assign \textit{random} HD vector codes to a set of "base" elements of the data (e.g.
+symbols), then impose \textit{a priori} rules for composing these into richer structures (e.g.
 words) \cite{Plate:1995, Kanerva:2009, Gayler:2004}.
-HDC's advantages are (1) random HD vectors have low overlap with high probability, enabling large element codebooks, (2) there is no hard limit on data complexity, and (3) encoding is often analytically tractable.
-While most systems would of course require some training on environment statistics, HDC provides a compelling blueprint for representational flexibility.
+HDC's advantages are (1) random HD vectors tend to have very low overlap, enabling large element codebooks, (2) there is no hard limit on data complexity, and (3) encoding is often analytically tractable.
+While most systems must of course learn some environment statistics, HDC provides a compelling blueprint for flexible representation.
 
-Most HDC schemes use codes in which relations among elements can be easily recovered, but only if one provides certain elements as cues (e.g.
+An ideal code should allow fast readout of both a composition's base elements and their relations, but most HDC schemes support one or the other.
+Sums of random HD vectors, for instance, have high overlap (low distance under most metrics) to their elements, enabling element readout via overlaps with the sum \cite{Bloom:1970, Anderson:1973, Plate:1994}; but cannot encode relations.
+HDC operations like circular convolution \cite{Plate:1995}, permutation \cite{Gayler:1998, Sahlgren:2008}, or matrix multiplication \cite{Gosmann:2019}, yield codes with cue-recoverable relations (cue A recovers B from ((A,B),(C,D))), but low overlap to the composition's base elements, precluding fast element readout without cues (but see \cite{Rachkovskij:2001}); this also conflicts with our desire that AB's code should resemble A's more than C's.
+Here we present a new HD composition operation inspired by chromosomal crossover, where two vectors $X$ and $X'$ are ordered or bound by masking out a fraction of $X$'s components and replacing them with those from $X'$; relational information is encoded in the context-dependent mask $\mu(X, X')$.
+"Crossover" codes thus highly overlap with their base elements, allowing fast recovery of both elements and relations, and decoding by greedy reconstruction.
+Further, they evenly distribute information and exhibit object-complexity-invariant statistics, ideal for robust intelligent systems.
+We describe crossover for sequences first, then generalize to binding and trees.
+
+% Most HDC schemes use codes in which relations among elements can be easily recovered, but only if one provides certain elements as cues (e.g.
 cue A recovers B from ((A,B),(C,D))).
@@
 Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states, $x_j \in G \equiv \{1,...,Z\}$, and let $d(X, X')$ be the Hamming distance from $X$ to $X'$.
-For a symbol sequence $Y = (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, we establish a mapping from $Y$ to $X^Y$ as follows:
+For a symbol sequence $Y \equiv Y_{1:L} \equiv (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, we establish a mapping from $Y$ to $X^Y$ as follows:
 
-First sample a codebook $C \equiv \{X^1, ..., X^M, X^*\}$, where $X^*$ is a "start code", and $x^j$ are i.i.d.
-across all components and codes and sampled uniformly from $G$.
-Next sample a mask function $\mu$, an $N \times M^{2N}$ matrix with i.i.d.
-real-valued elements from $\textrm{Uniform}(0, 1)$, which we reference as $\mu_j(X, X')$.
-In practice, columns of $\mu$ can be sampled "as needed" with a hash function.
+We first sample a codebook $C \equiv \{X^1, ..., X^M, X^*\}$, where $X^*$ is a "start code", and $x_j$ are i.i.d.
+across all components and symbols and sampled uniformly from $G$.
+Next, we sample a mask function $\mu$, an $N \times M^{2N}$ matrix with i.i.d.
+real-valued elements from $\textrm{Uniform}(0, 1)$, referenced as $\mu_j(X, X')$.
+In practice, columns of $\mu$ can be sampled "as needed" using a pseudorandom number generator.
 
-Using $C$ and $\mu$ we encode sequences as in Alg 1 and Fig 1a: symbol codes are recursively woven into the final code by replacing a fraction of the current code's components with those from the symbol code; which components are replaced encodes sequence order.
-We decrease the fraction as $1/(t+1)^\gamma$, for $\gamma \geq 1$ to prevent exponential decay of the first symbol's presence in the final code.
+Using $C$ and $\mu$ we encode sequences as in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": at each $t$ we mask out a fraction $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$.
+The mask selects components where $\mu_j(X^{Y_{1:t-1}}, X^{y_t}) < 1/(t+1)^\gamma$, so is unique for all $(X^{Y_{1:t-1}}, X^{y_t})$, yielding context-dependent mixing.
+$\gamma \geq 1$ prevents decay of early symbols in $X^Y$.
 
@@
 \EndFunction
-
-\State
 
@@
 
-For $\gamma = 1$, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$; $n^i$ is how often $i$ appears in $Y$.
-I.e.
-for large $Z$, mean symbol-sequence similarity (over $C, \mu$) scales with the symbol's count within the sequence (Fig 1b), with variance $\sim 1/N$.
-As exemplified by Fig 1c, crossover code components are uniform over $G$, i.i.d.
+For large $Z$ and $\gamma = 1$, mean symbol-sequence similarity (over $C, \mu$) scales with symbol $i$'s count, $n^i$, within the sequence (Figure 1b), with variance $\sim 1/N$.
+Specifically, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$.
+As exemplified by Figure 1c, crossover code components are uniform over $G$, i.i.d.
 for a given $Y$, and invariant to sequence statistics (e.g.
@@
 \right]$$
-For most $Y$, $X^Y$ overlaps little with $Y$'s anagrams, and more with correct than incorrect starting subsequences, e.g.
-codes for CI and CIC are closer to CICADA's code than codes for CD and CIA, respectively (Fig 1d,e).
-This is because different next-symbol candidates yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols.
-Consequently, $Y$ can greedily decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step the symbol maximizing overlap with $X^Y$ when woven into the current code, then weaving it in.
-Of note, decoding as such is noise-robust, and when multiple $Y$ are given via a prior $P(Y)$, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
-across $j$, yielding equivalent robustness to random and targeted attack, so decoding accuracy depends only on the number of corrupted components (Fig 1f).
+For most $Y$, $X^Y$ overlaps little with codes for $Y$'s anagrams, and more with correct than incorrect starting subsequences, e.g.
+CI's and CIC's codes are closer than CD's and CIA's codes, respectively, to CICADA's code (Figure 1d,e).
+This is because different next-symbols yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols.
+Consequently, $Y$ can "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step $t$ the symbol $i$ maximizing overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$.
+Decoding is robust to noise added to $X^Y$, and when a prior $P(Y)$ exists, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
+across $j$; robustness is thus equivalent to random and targeted attacks, so decoding accuracy depends only on the number of corrupted $x_j^Y$ (Figure 1f).
 
-For $Y$ with many symbol repeats distinct from $y_1$ (e.g.
+While crossover intrinsic supports sequences with repeated symbols (either consecutive or not), for $Y$ with many symbol repeats distinct from $y_1$ (e.g.
 REFEREE), $X^*$ crossed over with $X^{y_1}$ can be further from $X^Y$ than $X^*$ crossed with the repeated symbol, impairing decoding.
-Making $\gamma > 1$, however, fixes this by more heavily weighting early symbols (Fig 2a).
-Surprisingly, although this front-loaded weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared across many sequences (Fig 2b).
-In general, decoding accuracy decreases with L and increases with N, as the variance of $P(x_j^Y = x_j^Y')$, with $Y'$ a starting subsequence of $Y$ scales as $1/N$.
-Notably, decoding accuracy depends little of $M$, the dictionary size, due to the large expected distances between HD random vectors.
+Making $\gamma > 1$ fixes this by more heavily weighting early symbols (Figure 2a).
+Surprisingly, while this front-loaded weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared by many sequences (Figure 2b).
+In general, decoding accuracy decreases (increases) with $L$ ($N$), since $\textrm{Var}[P(x_j^{Y_{1:t}} = x_j^Y)] \sim 1/N$.
+Decoding accuracy is also robust to large $M$, the dictionary size, due at heart to the low expected overlap of HD random vectors.
 
@@
 While crossover with i.i.d.
-$\mu$ is robust to noise added to the final code (Fig 1f), it fails when noise is added during encoding (Fig 2c,d), as one altered component yields an entirely different mask.
-This is fixed by introducing correlations into $\mu_j(X, X')$ (here making $\mu$ a function of sparse components of $X, X'$) (Fig 2c,d).
-The tradeoff is decreased decoding accuracy for certain sequences, since correlated masks are less distinct, but this can be corrected via increased $N$.
+$\mu$ is robust to noise added to the final code (Figure 1f), it fails when noise is added during encoding (Figure 2c,d), as one altered component yields an entirely different mask.
+This is fixed by introducing correlations into $\mu_j(X, X')$ (here by making $\mu$ a function of sparse components of $X, X'$) (Figure 2c,d).
+The cost of this is decreased decoding accuracy for certain sequences, since correlated masks are less distinct, but this can be corrected by increasing $N$.
 
-Beyond sequences, crossover naturally generalizes to nested structures, or trees.
-Key to encoding trees is \textit{binding} elements together into groups that can themselves be bound, yet without forgetting initial group identities, e.g.
-((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)).
-In crossover, $k$ elements $i_0, ..., i_k$ can be bound into a group by using a context-dependent mask function $\mu(X^{i_0}, ....
-X^{i_k})$ (which can commute if within-group order does not matter).
+Beyond sequences, crossover naturally generalizes to \textit{k}-ary trees.
+Key to encoding trees is \textit{binding} elements into groups that can themselves be bound, yet without forgetting initial group identities: ((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)).
+In crossover, $k$ elements $i_0, ..., i_k$ can be bound into a group by using a context-dependent mask function $\mu_j(X^{i_0}, ....
+X^{i_k})$ (optionally commuting).
 The vector for the group takes components from $X^{i_0}$ where $0 < \mu < 1/k$, from $X^{i_1}$ where $1/k < \mu < 2/k$, etc.
-This vector can then be bound itself in the exact same manner (Fig 3a), allowing one to recursively encode a tree of arbitrary depth.
-The result is a vector overlapping more with elements within vs not in the tree, and with groups within vs not in tree, even if the latter contain same base elements (Fig 3b).
-Generally, a tree encoding by crossover will tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
-Similar to sequence encoding, this means that both base elements and their recursive relationships can be extracted via their overlap with the crossover code, and that decoding of the full tree can occur through a recursive, but greedy reconstruction.
+This can then be bound itself in the exact same manner (Figure 3a), allowing recursive encoding of an arbitrary tree.
+The result is a vector overlapping more with elements in vs not in the tree, and with groups in vs not in tree, even if the latter contain the same base elements (Figure 3b).
+A generic tree encoding will tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
+Similar to sequence encoding, this means both base elements and their recursive relations can be extracted via their overlap with the final crossover code, and that decoding of the full tree can occur through a recursive, greedy reconstruction.
 
@@
   \includegraphics{3_trees.pdf}
-  \caption{Tree/binding encoding via crossover ($N=1024, Z=8, M=26$). (a) Example crossover encoding of ((A,B), (C,D), (E,F)). (b) Distances of final code to element and bound-pair codes.}
+  \caption{Tree/binding encoding via crossover ($N=1024, Z=8, M=26$). (a) Example crossover encoding of ((A,B), (C,D), (E,F)). (b) Distances of final code to element and pair/group codes.}
 \end{figure}
@@
 
-While intelligent systems must learn environment statistics to operate successfully, it can useful for certain computational features to be built in \cite{Zador:2019}.
-Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects/events if they are made from familiar elements: an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only learned the isolated scene elements.
+While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built in \cite{Zador:2019}.
+Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only learned the isolated scene elements.
 While many trained systems, e.g.
@@
 Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
-Basing training on convolutional binding, for instance, allowed an artificial network to learn knowledge graphs more efficiently and scalably than other methods \cite{Nickel:2016}.
-It stands to reason that identifying robust HDC codes could provide a useful backbone to improve artificial systems, as well as for directing our investigation of neural computation in biology.
+Basing training on convolutional binding, for instance, allowed an artificial network to efficiently and scalably learn knowledge graphs and out-perform state-of-the-art methods at link prediction \cite{Nickel:2016}.
+It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, as well as to guiding research in neural computation in biology.
 
 "Crossover" shares advantages with existing HDC schemes but is distinct in key ways also.
-Like other schemes, crossover preserves relational information among elements, which is readily recoverable from the final code \cite{Kanerva:2009}.
-Further, relation codes can be reused as elements themselves, allowing arbitrary deep recursive composition \cite{Kanerva:2009}; in crossover one can add CAFE to the dictionary (perhaps with minor noise added), enabling encoding of longer sequences made from shorter ones.
-To encode relations, however, previous schemes have relied on circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1994}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yields relation codes that little overlap their elements.
-While these distinguish different compositions of the same elements and allow cue-based recall, base elements cannot be identified directly from their overlap with the relation code, and instead require additional storage.
-An exception is "context-dependent thinning" (CDT), which stores relations via the union of the elements, then "stamping" the code with a context-dependent set of zeros \cite{Rachkovskij:2001}; while the stamp shape retains relational information, however, information beneath its imprint is discarded, and it is not clear how CDT could construct recursive sequence codes without early elements decaying exponentially.
+Like other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrary deep composition.
+For instance, in crossover one can add CAFE's or other words' codes to the dictionary (perhaps with minor noise added), enabling encoding of long sequences made from short ones.
+To encode relations, previous schemes have used circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1994}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yield relation codes that little overlap their elements.
+While these distinguish different compositions of the same elements and allow cue-based recall, base elements cannot be read out directly from their overlap with the relation code, and the entire dictionary must be used (instead of just the relevant symbols) when reading out relations, potentially greatly slowing decoding for large $M$.
+An exception is "context-dependent thinning" (CDT), which stores relations via the union of the elements, then "stamps" the code with a context-dependent set of zeros \cite{Rachkovskij:2001}; while the stamp shape retains relational information without erasing element overlaps, however, information beneath its imprint is discarded, and it is not clear how CDT could construct recursive sequence codes without early elements decaying exponentially.
 Crossover overcomes these limitations, as both elements and their relations can be read out directly, and in sequences no element overlaps decay exponentially.
@@
 Crossover is also conceptually distinct from its predecessors since one need not define arithmetical operations over code components.
-Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Kanerva:1994, Gayler:1998, Rachkovskij:2001, Sahlgren:2008, Gosmann:2019}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
+Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Gayler:1998, Sahlgren:2008, Gosmann:2019, Rachkovskij:2001, Kanerva:1994}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
 Crossover thus outlines a distributed computation substrate that can exist outside vector spaces, and instead within more generic metric spaces.
 This is relevant to biological implementation, as it allows one to envision $x_j$ not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank.
-For instance, each component of a crossover code could correspond to the attractor state of a 100-neuron Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree.
+For instance, each component of a crossover code could correspond to the attractor state of a small Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree.
 Such could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code overlap statistics to be usefully read out and manipulated by downstream and recurrent computations.

### 2019-09-13 ms_neurips_2019/ms_neurips_2019.tex
Do final revisions before submission.

@@
 % ready for submission
-% \usepackage{neurips_2019}
+\usepackage{neurips_2019}
 
@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-     \usepackage[final]{neurips_2019}
+%     \usepackage[final]{neurips_2019}
 
@@
 HDC operations like circular convolution \cite{Plate:1995}, permutation \cite{Gayler:1998, Sahlgren:2008}, or matrix multiplication \cite{Gosmann:2019}, yield codes with cue-recoverable relations (cue A recovers B from ((A,B),(C,D))), but low overlap to the composition's base elements, precluding fast element readout without cues (but see \cite{Rachkovskij:2001}); this also conflicts with our desire that AB's code should resemble A's more than C's.
-Here we present a new HD composition operation inspired by chromosomal crossover, where two vectors $X$ and $X'$ are ordered or bound by masking out a fraction of $X$'s components and replacing them with those from $X'$; relational information is encoded in the context-dependent mask $\mu(X, X')$.
+Here we present a new HD composition operation inspired by chromosomal crossover, where two vectors $X$ and $X'$ are ordered or bound by masking out a fraction of $X$'s components and replacing them with those from $X'$; relational information is encoded in the context-dependent mask $\mu_:(X, X')$.
 "Crossover" codes thus highly overlap with their base elements, allowing fast recovery of both elements and relations, and decoding by greedy reconstruction.
@@
 
-Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states, $x_j \in G \equiv \{1,...,Z\}$, and let $d(X, X')$ be the Hamming distance from $X$ to $X'$.
-For a symbol sequence $Y \equiv Y_{1:L} \equiv (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, we establish a mapping from $Y$ to $X^Y$ as follows:
+Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states: $x_j \in G \equiv \{1,...,Z\}$; and let $d(X, X')$ be the Hamming distance from $X$ to $X'$.
+For a symbol sequence $Y \equiv Y_{1:L} \equiv (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, a dictionary of $M$ symbols, we map $Y$ to $X^Y$ as follows:
 
@@
 
-Using $C$ and $\mu$ we encode sequences as in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": at each $t$ we mask out a fraction $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$.
+Using $C$ and $\mu$ we encode sequences as shown in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": at each $t$ we mask out a fraction $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$.
 The mask selects components where $\mu_j(X^{Y_{1:t-1}}, X^{y_t}) < 1/(t+1)^\gamma$, so is unique for all $(X^{Y_{1:t-1}}, X^{y_t})$, yielding context-dependent mixing.
@@
 
-\State $mask \gets (\mu(X, X') < 1/(t+1)^\gamma) \quad$  \# context-dependent mask with about $1/(t+1)^\gamma$ 1's
+\State $mask \gets (\mu_:(X, X') < 1/(t+1)^\gamma) \quad$  \# context-dependent mask with about $1/(t+1)^\gamma$ 1's
 \State $X'' \gets X$
@@
   \caption{Sequence encoding via crossover ($N=1024, Z=8, M=26, \gamma=1$).
-(a) Example encoding of CAFE; colors are states, inscribed symbols are visual aids, not used by the algorithm.
+(a) Example encoding of CAFE; colors are states; inscribed symbols are visual aids, not used by the algorithm.
 (b) Sequence-symbol code distances.
@@
 (f) Decoding vs code corruption.
-"x"s indicate theoretical predictions and bars average simulation results.
-All plots here and elsewhere were made via 100 samples of $C, \mu$.
-"*" prefixes indicate sequences.}
+All plots here/elsewhere were made via 100 $C, \mu$ samples.
+"x"s denote exact calculations and bars average simulation results.
+"*" prefixes distinguish sequences from symbols.}
   \label{fig:1}
@@
 
-For large $Z$ and $\gamma = 1$, mean symbol-sequence similarity (over $C, \mu$) scales with symbol $i$'s count, $n^i$, within the sequence (Figure 1b), with variance $\sim 1/N$.
+For large $Z$ and $\gamma = 1$, mean similarity ($1-d$) of a symbol $i$ to the final sequence scales with symbol $i$'s count, $n^i$, within the sequence (Figure 1b), with variance $\sim 1/N$.
 Specifically, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$.
@@
 $$+\frac{1}{L+1}\frac{1}{L'+1}\left[
-\mathbf{n}^T_{u}\mathbf{n}_{v'} + \left((t+1)(L'-t) - \mathbf{n}^T_{u}\mathbf{n}_{v'} \right)\frac{1}{Z}
-+ \mathbf{n}^T_{v}\mathbf{n}_{u} + \left((L-t)(t+1) - \mathbf{n}^T_{v}\mathbf{n}_{u} \right)\frac{1}{Z}
+\mathbf{n}^T_{u}(\mathbf{n}_{v} + \mathbf{n}_{v'}) + \left((t+1)(L'+L-2t) - \mathbf{n}^T_{u}(\mathbf{n}_{v} + \mathbf{n}_{v'})
+ \right)\frac{1}{Z}
 \right]$$
@@
 This is because different next-symbols yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols.
-Consequently, $Y$ can "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step $t$ the symbol $i$ maximizing overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$.
+Consequently, $Y$ can be "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step $t$ the symbol $i$ that maximizes overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$.
 Decoding is robust to noise added to $X^Y$, and when a prior $P(Y)$ exists, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
-across $j$; robustness is thus equivalent to random and targeted attacks, so decoding accuracy depends only on the number of corrupted $x_j^Y$ (Figure 1f).
+across $j$; robustness is thus equivalent for random and targeted attacks, with decoding accuracy depending only on the number of corrupted $x_j^Y$ (Figure 1f).
 
-While crossover intrinsic supports sequences with repeated symbols (either consecutive or not), for $Y$ with many symbol repeats distinct from $y_1$ (e.g.
+While crossover intrinsically supports sequences with repeated symbols (either consecutive or not), for $Y$ with many symbol repeats distinct from $y_1$ (e.g.
 REFEREE), $X^*$ crossed over with $X^{y_1}$ can be further from $X^Y$ than $X^*$ crossed with the repeated symbol, impairing decoding.
@@
 and correlated $\mu$.
-(d) Decoding of *CICADA for i.i.d., correlated $\mu$, and correlated $\mu$ with increased $N$.}
+(d) Decoding of *CICADA for i.i.d., correlated $\mu$, and correlated $\mu$ with quadrupled $N$.}
 \end{figure}
@@
 $\mu$ is robust to noise added to the final code (Figure 1f), it fails when noise is added during encoding (Figure 2c,d), as one altered component yields an entirely different mask.
-This is fixed by introducing correlations into $\mu_j(X, X')$ (here by making $\mu$ a function of sparse components of $X, X'$) (Figure 2c,d).
-The cost of this is decreased decoding accuracy for certain sequences, since correlated masks are less distinct, but this can be corrected by increasing $N$.
+This is fixed by introducing correlations into $\mu_j(X, X')$ (i.e.
+making $\mu$ a smooth function of $X, X'$ [many methods work]).
+The cost of this is decreased decoding accuracy for certain sequences, since correlated are less distinct than i.i.d.
+masks, but this can be corrected by increasing $N$ (Figure 2c,d).
 
@@
 Key to encoding trees is \textit{binding} elements into groups that can themselves be bound, yet without forgetting initial group identities: ((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)).
-In crossover, $k$ elements $i_0, ..., i_k$ can be bound into a group by using a context-dependent mask function $\mu_j(X^{i_0}, ....
-X^{i_k})$ (optionally commuting).
-The vector for the group takes components from $X^{i_0}$ where $0 < \mu < 1/k$, from $X^{i_1}$ where $1/k < \mu < 2/k$, etc.
+In crossover, $k$ elements $i_1, ..., i_k$ can be bound into a group using a context-dependent mask function $\mu_j(X^{i_1}, ....
+X^{i_k})$ (which can optionally commute).
+The vector for the group takes components from $X^{i_1}$ where $0 < \mu < 1/k$, from $X^{i_2}$ where $1/k < \mu < 2/k$, etc.
 This can then be bound itself in the exact same manner (Figure 3a), allowing recursive encoding of an arbitrary tree.
 The result is a vector overlapping more with elements in vs not in the tree, and with groups in vs not in tree, even if the latter contain the same base elements (Figure 3b).
-A generic tree encoding will tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
-Similar to sequence encoding, this means both base elements and their recursive relations can be extracted via their overlap with the final crossover code, and that decoding of the full tree can occur through a recursive, greedy reconstruction.
+A generic tree encoding will also tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
+Similar to sequence encoding, this means both base elements and their recursive relations can be extracted via their overlap with the final code, and that decoding of the full tree can occur through recursive but greedy reconstruction.
 
@@
 While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built in \cite{Zador:2019}.
-Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only learned the isolated scene elements.
+Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with the isolated scene elements.
 While many trained systems, e.g.
-translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}).
+translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}).
 Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
 Basing training on convolutional binding, for instance, allowed an artificial network to efficiently and scalably learn knowledge graphs and out-perform state-of-the-art methods at link prediction \cite{Nickel:2016}.
-It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, as well as to guiding research in neural computation in biology.
+It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, as well as to guiding investigation of neural computation in biology.
 
 "Crossover" shares advantages with existing HDC schemes but is distinct in key ways also.
-Like other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrary deep composition.
+As in other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrary deep composition.
 For instance, in crossover one can add CAFE's or other words' codes to the dictionary (perhaps with minor noise added), enabling encoding of long sequences made from short ones.
@@
 Further, crossover yields i.i.d.
-information across code components and code statistics invariant to object complexity, suggesting it may be a useful representation for a robust intelligent system.
+information across code components, and code statistics are invariant to object complexity, suggesting it may be a useful representation for a robust intelligent system.
 
 Crossover is also conceptually distinct from its predecessors since one need not define arithmetical operations over code components.
-Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Gayler:1998, Sahlgren:2008, Gosmann:2019, Rachkovskij:2001, Kanerva:1994}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
+Whereas previous HDC algorithms require \textit{summing} components, either directly or through convolution \cite{Plate:1995, Gayler:1998, Sahlgren:2008, Gosmann:2019, Rachkovskij:2001, Kanerva:1994}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation.
 Crossover thus outlines a distributed computation substrate that can exist outside vector spaces, and instead within more generic metric spaces.
@@
 For instance, each component of a crossover code could correspond to the attractor state of a small Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree.
-Such could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code overlap statistics to be usefully read out and manipulated by downstream and recurrent computations.
+This could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code-overlap statistics to be usefully read out and manipulated by further downstream and recurrent computations.
 
-\subsubsection*{Acknowledgments}
+% \subsubsection*{Acknowledgments}
 
-Use unnumbered third level headings for the acknowledgments. All acknowledgments
-go at the end of the paper. Do not include acknowledgments in the anonymized
-submission, only in the final paper.
+% Use unnumbered third level headings for the acknowledgments. All acknowledgments
+% go at the end of the paper. Do not include acknowledgments in the anonymized
+% submission, only in the final paper.
 

### 2019-11-08 ms_neurips_2019/ms_neurips_2019.tex
Add appendix.

@@
 % ready for submission
-\usepackage{neurips_2019}
+% \usepackage{neurips_2019}
 
@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2019}
+    \usepackage[final]{neurips_2019}
 
@@
 \usepackage[noend]{algpseudocode}
+\usepackage[title]{appendix}
 % \usepackage[superscript,biblabel]{cite}
@@
 Inspired by chromosomal crossover, random HD vectors are recursively interwoven, with a fraction of one vector's components masked out and replaced by those from another using a context-dependent mask.
-Unlike most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes, allowing fast element readout without sacrificing relational information, and decoding by greedy reconstruction.
+Unlike most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
 Crossover is mathematically tractable and has several properties desirable for robust, flexible representation.
@@
 Ideally, similar objects should have similar codes, new objects should not require rearranging existing codes, and there should be no limit on object complexity.
-Distributing codes across high-dimensional (HD) vectors yields potential robustness, flexibility, and generalization \cite{Hinton:1984, Mikolov:2013}; but how does one store compositional relations, like order or binding in such a code, e.g.
+Distributing codes across high-dimensional (HD) vectors offers a solution with potential robustness, flexibility, and generalization \cite{Hinton:1984, Mikolov:2013}; but how does one store compositional relations, like order or binding, in such a code, e.g.
 to distinguish AB vs BA or ((A,B), (C,D)) vs ((A,D), (C,B)), respectively?
@@
 
-\section{Results}
+\section{Coding scheme}
 
@@
 across all components and symbols and sampled uniformly from $G$.
-Next, we sample a mask function $\mu$, an $N \times M^{2N}$ matrix with i.i.d.
+Next, we sample a mask function $\mu$, an $N \times Z^{2N}$ matrix with i.i.d.
 real-valued elements from $\textrm{Uniform}(0, 1)$, referenced as $\mu_j(X, X')$.
@@
 
-Using $C$ and $\mu$ we encode sequences as shown in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": at each $t$ we mask out a fraction $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$.
+Using $C$ and $\mu$ we encode sequences as shown in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": for the $t$'th element we mask out $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$.
 The mask selects components where $\mu_j(X^{Y_{1:t-1}}, X^{y_t}) < 1/(t+1)^\gamma$, so is unique for all $(X^{Y_{1:t-1}}, X^{y_t})$, yielding context-dependent mixing.
@@
 
-For large $Z$ and $\gamma = 1$, mean similarity ($1-d$) of a symbol $i$ to the final sequence scales with symbol $i$'s count, $n^i$, within the sequence (Figure 1b), with variance $\sim 1/N$.
-Specifically, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$.
+For large $Z$ and $\gamma = 1$, mean similarity ($1-d$) of symbol code $X^i$ to the final sequence code $X^Y$ scales with $i$'s count within the sequence, $n^i$ (Figure 1b), with variance $\sim 1/N$.
+Precisely, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q \equiv P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$.
 As exemplified by Figure 1c, crossover code components are uniform over $G$, i.i.d.
@@
 The proof follows from the i.i.d.
-nature of $C, \mu$ and the interchangeability of component states, and suggests crossover as a robust, scale-invariant encoding.
+nature of $C, \mu$ and the interchangeability of component states, suggesting crossover as a robust, scale-invariant encoding.
 
@@
 This is because different next-symbols yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols.
-Consequently, $Y$ can be "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step $t$ the symbol $i$ that maximizes overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$.
+Consequently, $Y$ can be "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at the $t$-th rebuild step the symbol $i$ that maximizes overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$.
 Decoding is robust to noise added to $X^Y$, and when a prior $P(Y)$ exists, the mutual information between $x_j^Y$ and $Y$ is i.i.d.
@@
 $\gamma$ for several sequences.
-(c) Decoding of *CAFE vs number of corrupted components at encoding step for i.i.d.
+(c) Decoding of *CAFE vs number of corrupted components at each encoding step for i.i.d.
 and correlated $\mu$.
-(d) Decoding of *CICADA for i.i.d., correlated $\mu$, and correlated $\mu$ with quadrupled $N$.}
+(d) Decoding of *CICADA for i.i.d.
+$\mu$, correlated $\mu$, and correlated $\mu$ with quadrupled $N$.}
 \end{figure}
@@
 Key to encoding trees is \textit{binding} elements into groups that can themselves be bound, yet without forgetting initial group identities: ((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)).
-In crossover, $k$ elements $i_1, ..., i_k$ can be bound into a group using a context-dependent mask function $\mu_j(X^{i_1}, ....
-X^{i_k})$ (which can optionally commute).
+In crossover, $k$ elements $i_1, ..., i_k$ can be bound into a group using a context-dependent mask function $\mu_j(X^{i_1}, ..., X^{i_k})$ (which can optionally commute).
 The vector for the group takes components from $X^{i_1}$ where $0 < \mu < 1/k$, from $X^{i_2}$ where $1/k < \mu < 2/k$, etc.
 This can then be bound itself in the exact same manner (Figure 3a), allowing recursive encoding of an arbitrary tree.
-The result is a vector overlapping more with elements in vs not in the tree, and with groups in vs not in tree, even if the latter contain the same base elements (Figure 3b).
+The result is a vector overlapping more with elements in vs not in the tree, and with groups in vs not in the tree, even if the latter contain the same base elements (Figure 3b).
 A generic tree encoding will also tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity.
@@
 
-While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built in \cite{Zador:2019}.
+While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built-in \cite{Zador:2019}.
 Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with the isolated scene elements.
@@
 translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}).
-Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
+Alternatively, one can design HDC algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
 Basing training on convolutional binding, for instance, allowed an artificial network to efficiently and scalably learn knowledge graphs and out-perform state-of-the-art methods at link prediction \cite{Nickel:2016}.
@@
 "Crossover" shares advantages with existing HDC schemes but is distinct in key ways also.
-As in other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrary deep composition.
+As in other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrarily deep composition.
 For instance, in crossover one can add CAFE's or other words' codes to the dictionary (perhaps with minor noise added), enabling encoding of long sequences made from short ones.
@@
 
-% \subsubsection*{Acknowledgments}
+\subsubsection*{Acknowledgments}
+
+I acknowledge...
 
@@
 \bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
+
+\begin{appendices}
+\section{Properties of crossover codes}
+
+\subsection{Proof that sequence codes have i.i.d. components}
+
+Denoting $Y_{1:t} \equiv (y_1, ..., y_t)$ with $Y_{1:L} = Y$, and $C_j \equiv (x^1_j, ..., x^M_j, x^*_j)$, we show that $x_j^{Y_{1:t}}$ are independent across $j$ for all $t$.
+
+We will use the fact that if $A_1, ..., A_N$ are independent, then so are $f(A_1), ..., f(A_N)$.
+In particular, define $\xi^t_j \equiv \{x_j^{Y_{1:t-1}}, ..., x_j^{Y_{1:0}}, C_j, \mu(j, X^*, X^{y_1}), \mu(j, X^{Y_{1:1}}, X^{y_2}), ..., \mu(j, X^{Y_{1:t-1}}, X^{y_t})\}$.
+Then $x_j^{Y_{1:t}}$ is a strict function of $\xi^t_j$ so it suffices to show $\xi^t_j$ are independent across $j$ for all $t$.
+
+We proceed by induction:
+
+\textbf{Base case:} $\xi^1_j = \{x^{Y_{1:0}}_j, C_j, \mu(j, X^{Y_{1:0}}, X^{y_1})\} = \{x^*_j, C_j, \mu(j, X^*, X^{y_1})\}$, which are all i.i.d., so $\xi^1_j$ are independent across $j$.
+
+\textbf{Inductive step:} If $\xi^t_j$ are independent across $j$ then so are $\{x_j^{Y_{1:t}}\} \cup \xi^t_j$, since this is a deterministic function of $\xi^t_j$.
+Further, $\{x_j^{Y_{1:t}}\} \cup \xi^t_j \cup \{\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})\}$ must also be independent across $j$.
+To see this, consider the following two cases.
+
+First, most likely is that $\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})$ has never been used in the construction of $X^{Y_{1:t-1}}$, so it is independent from $\{x_j^{Y_{1:t}}\} \cup \xi^t_j$ while also independent across $j$.
+Therefore, their union is independent across $j$.
+
+In the second, less likely case, $\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})$ \textit{was} used to construct $X^{Y_{1:t-1}}$ at some point in the past, i.e.
+$(X^{Y_{1:t}}, X^{y_{t+1}}) = (X^{Y_{1:t'}}, X^{y_{t'+1}})$ for some $t' < t$.
+In this case, however, the term $\mu(j, X^{Y_{1:t}}, X^{y_{t+1}}) = \mu(j, X^{Y_{1:t'}}, X^{y_{t'+1}})$ is already included in $\xi^t_j$, and so the independence across $j$ is not affected because no new random variables are added.
+Thus, $\{x_j^{Y_{1:t}}\} \cup \xi^t_j \cup \{\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})\}$ is independent across $j$.
+
+Since $\{x_j^{Y_{1:t}}\} \cup \xi^t_j \cup \{\mu(j, X^{Y_{1:t}}, X^{y_{t+1}})\} = \xi^{t+1}_j$, we have that $\xi^t_j$ being independent across $j$ implies that $\xi^{t+1}_j$ is independent across $j$, completing the inductive step.
+
+Therefore $x_j^{Y_{1:t}}$ is independent across $j$ for all $t$, in other words $P(X^Y|Y) = \prod\limits_{j=1}^N P(x^Y_j|Y)$.
+
+The equality $P(x^Y_j|Y) = P(x^Y_{j'}|Y) \quad \forall j, j'$ arises because no component is treated differently from any other in our algorithm, so symmetry implies they can have no statistical differences.
+
+\subsection{Proof that component distributions are uniform over component states}
+
+$P(x^Y_j = g|Y) = P(x^Y_j = g'|Y)$ for $g \neq g'$ arises because all $g$ in our model are treated identically, so symmetry implies any functions of them can have no statistical differences.
+Since these must add to $1$ when summed over $g \in G$, and $|G| = Z$, we must also have $P(x^Y_j = g|Y) = 1/Z.$
+
+This also implies that the marginal statistics of crossover codes do not depend on $Y$ or any feature of $Y$ e.g.
+sequence length.
+This distinguishes crossover, for instance, from representations that would grow denser or sparser with the length of the sequence they represent.
+
+\subsection{Proof that sequence information is i.i.d. across components}
+
+Given $C$, $\mu$, and a prior distribution over sequences $P_Y(Y)$ with entropy $H[Y]$ over $N_Y$ different sequences in total, define the information $I_j | C, \mu$ in component $j$ as the mutual information between $Y$ and $x_j$, i.e.
+
+$$I_j | C, \mu \equiv MI[Y|x_j; C, \mu] = H[Y] - H[Y|x_j; C, \mu] = $$
+$$H[Y] - \sum_g H[Y|x_j = g; C, \mu] P_Y(x_j = g| C, \mu).$$
+
+The components for one sequence $Y$ are i.i.d.
+Therefore, while the $x^Y_j$ and $x^{Y'}_j$ might be dependent (e.g.
+if $Y$ and $Y'$ are similar), there can be no dependencies between $x^Y_j$ and $x^{Y'}_{j'}$ for $j \neq j'$.
+Therefore $\{x^{Y_1}_j, ..., x^{Y_{N_Y}}_j\}$ is independent across $j$.
+Since $I_j$ is a deterministic function of $\{x^{Y_1}_j, ..., x^{Y_{N_Y}}_j\}$ (quantifying how many possible sequences get eliminated upon knowing the value of $x^Y_j$), $I_j$ is independent across $j$ also.
+Since no $j$ is treated specially in our algorithm, $I_j$ must also be identically distributed, so given an initial prior over sequences to encode $P_Y(Y)$, $I_j$ is i.i.d.
+
+\section{Symbol/sequence code overlap derivations}
+
+Since $x^Y_j$ are i.i.d., the similarity or overlap between two sequence codes  (1 minus the Hamming distance) $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$.
+The randomness giving rise to the distribution is over samples of $C$ and $\mu$.
+Thus, $\textrm{E}[s(X^Y, X^{Y'})] = q$ and $\textrm{Var}[s(X^Y, X^{Y'})] = q(1-q)/N$, and it suffices to only compute $q$.
+Similar results hold for the similarity of a sequence and symbol code, $s(X^Y, X^i)$.
+For conciseness, we drop the $j$ and write $x \equiv x_j$ from now on.
+
+Our goal is to compute $q$ for an arbitrary sequence $Y$ and symbol $i$ and for two arbitrary sequences $Y$ and $Y'$, since this gives us the full distribution of similarity values over samples of $C$ and $\mu$.
+
+\subsection{Useful quantities}
+
+We first compute a few quantities to keep calculations more concise.
+
+1.
+$f^{\gamma,L}_t \equiv P(x^Y \leftarrow y_t)$ is the probability that $x^Y$ was taken from the $t$-th element of $Y$.
+This is $1/(t+1)^\gamma$ times the probability the component was not overwritten between $t+1$ and $L$ (including occasions where it was overwritten by the same state).
+We then have:
+
+$$f^{\gamma,L}_t \equiv P(x^Y \leftarrow y_t) = \frac{1}{(t+1)^\gamma}\prod\limits_{t' = t+1}^L\left(1 - \frac{1}{(t'+1)^\gamma}\right).$$
+
+Note that this is not a function of $x^Y$. Also note that $f^{1,L}_t = 1/(L+1)$, which is not a function of $t$.
+
+2.
+$P(x^Y \leftarrow i)$ is the probability that $x^Y$ was taken from symbol $i$, which will depend on how many times $i$ was present in $Y$.
+We compute this by summing the probabilities of all ways this could have occurred.
+
+Suppose $y_t = i$ for $t_1 < ...
+< t_{n^i}$, with $\sum_i n^i = L$.
+We can divide the combinatorial ways $x^Y$ could have ended up being taken from $i$ into $n^i$ cases.
+First, $x^Y$ could have been taken from $y_{t_1}$ then never overwritten, whose probability is $f^{\gamma,L}_{t_1}$.
+Next, it could have been taken from $y_{t_2}$ then never overwritten (regardless of whether it was taken from $y_{t_1}$), whose probability is $f^{\gamma,L}_{t_2}$.
+And so forth until $t_{n^i}$.
+These $n^i$ cases are mutually non-overlapping and cover all possible ways that $x^Y$ could have been taken from $i$.
+Thus
+
+$$P(x^Y \leftarrow i) = f^{\gamma,L}_{t_1} + ... + f^{\gamma,L}_{t_{n^i}}$$.
+
+When $\gamma = 1$, $P(x^Y \leftarrow i) = n^i/(L+1)$.
+
+3. $P(x^Y \leftarrow Y_{1:t})$ is the probability that $x^Y$ was not overwritten from $t+1$ to $L$. This is just
+
+$$P(x^Y \leftarrow Y_{1:t}) = \prod\limits_{t' = t+1}^L \left(1 - \frac{1}{(t'+1)^\gamma}\right) = (t+1)^\gamma f^{\gamma,L}_t.$$
+
+When $\gamma = 1$, $P(x^Y \leftarrow Y_{1:t}) = (t+1)/(L+1)$.
+
+4.
+$P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t})$ is the probability that $x^Y$ came from $i$ given that it did not come from $Y_{1:t}$ (i.e.
+given that it was overwritten at some point from $t+1$ to $L$).
+Suppose $y_s = i$ for $t < s_1 < ...
+< s_{n^i_v}$ where $n^i_v$ is the number of times symbol $i$ appears in $Y_{t+1:L} \equiv (y_{t+1}, ..., y_L)$, with $\sum\limits_i n^i_v = L-t$.
+
+We compute $P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t})$ by summing over the probabilities of all the ways this could have occurred. Similar to the calculation for $P(x^Y \leftarrow i)$, we have
+
+$$P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t}) = P(x^Y \leftarrow y_{s_1}| x^Y \not \leftarrow Y_{1:t}) + ... + P(x^Y \leftarrow y_{s_{n^i_v}}| x^Y \not \leftarrow Y_{1:t}).$$
+
+By Bayes' Rule:
+
+$$P(x^Y \leftarrow y_{s_k}| x^Y \not \leftarrow Y_{1:t}) = \frac{P(x^Y \not \leftarrow Y_{1:t}| x^Y \leftarrow y_{s_k})P(x^Y \leftarrow y_{s_k})}{P(x^Y \not\leftarrow Y_{1:t})} = \frac{P(x^Y \leftarrow y_{s_k})}{P(x^Y \not\leftarrow Y_{1:t})}$$
+
+since if $x^Y$ came from $y_{s_k}$, which is in the part of $Y$ after $t$, then $x^Y$ could have not have come from $Y_{1:t}$.
+We have already computed the numerator and denominator (which is just $1 - P(x^Y \leftarrow Y_{1:t}))$, so
+
+$$P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t}) = \frac{1}{P(x^Y \not\leftarrow Y_{1:t})}\left( f^{\gamma,L}_{s_1} + ... + f^{\gamma,L}_{s_{n^i_v}} \right).$$
+
+When $\gamma = 1$ this reduces to $P(x^Y \leftarrow i|x^Y \not \leftarrow Y_{1:t}) = n^i_v/(L-t)$, i.e. it is proportional to how many times $i$ appears in the second part of the sequence $Y_{t+1:L}$.
+
+We now calculate how similar a sequence code is to any symbol code, as well as to other sequence codes.
+
+\subsection{Derivation: Similarity between symbol and sequence codes}
+
+$$P(x^Y = x^i) = P(x^Y \leftarrow i)P(x^Y = x^i|x^Y \leftarrow i) + P(x^Y \not \leftarrow i)P(x^Y = x^i|x^Y \not \leftarrow i)
+$$
+$$= P(x^Y \leftarrow i) + P(x^Y \not \leftarrow i)\frac{1}{Z}$$
+
+$$= \sum\limits_{t \in (t_1, ..., t_{n^i})}f^{\gamma,L}_t + \left(1 - \sum\limits_{t \in (t_1, ..., t_{n^i})}f^{\gamma,L}_t\right)\frac{1}{Z}
+= \sum\limits_{t \in (t_1, ..., t_{n^i})}\left(1 - \frac{1}{Z}\right)f^{\gamma,L}_t + \frac{1}{Z}$$
+
+When $\gamma = 1$, $f_t^{1,L} = 1/(L+1)$ so $P(x^Y = x_i)$ simplifies to $n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$, which tends to  $n^i/(L+1)$ for large $Z$.
+
+\subsection{Derivation: Similarity between codes for two sequences}
+
+Let $u^i$ index the times $i$ appears in $Y_{1:t}$, $v^i$ the times $i$ appears in $Y_{t+1:L}$, and $(v')^i$ the times $i$ appears in $Y'_{t+1:L'}$.
+Let $n^i_u$ count how many times $i$ appears in $Y_{1:t}$, $n^i_v$ the number of times $i$ appears in $Y_{t+1:L}$, and $n^i_{v'}$ the number of times $i$ appears in $Y'_{t+1:L'}$.
+We also write these in vector format as $\mathbf{n}_u$, $\mathbf{n}_v$, and $\mathbf{n}_{v'}$, respectively, where $\mathbf{n}_u$, $\mathbf{n}_v$, and $\mathbf{n}_{v'}$ are length-$M$ vectors, with one component per symbol in the dictionary $D$.
+
+Since $Y_{1:t} = Y'_{1:t}$,  first note that $X^{Y_{1:t}} = X^{Y'_{1:t}}$, so $x^{Y_{1:t}} = x^{Y'_{1:t}}$. We wish to find $P(x^Y = x^{Y'})$.
+
+The event $x^Y = x^{Y'}$ can occur in 4 distinct ways, whose probabilities sum to $P(x^Y = x^{Y'})$. At present we assume a negligible chance two sequence or symbol codes are exactly identical.
+
+\textbf{Case 1}: $(x^Y \leftarrow Y_{1:t}) \land (x^{Y'} \leftarrow Y'_{1:t})$, i.e. neither $x$ is overwritten at $t+1$ or later. These events are independent, so
+
+$$P(\textrm{Case 1}) = P(x^Y \leftarrow Y_{1:t})P(x^{Y'} \leftarrow Y'_{1:t}) = (t+1)^\gamma f^{\gamma,L}_t(t+1)^\gamma f^{\gamma,L'}_t$$
+
+When $\gamma = 1$, then:
+
+$$P(\textrm{Case 1}) = \frac{t+1}{L+1}\frac{t+1}{L'+1}$$.
+
+\textbf{Case 2}: $(x^Y \not\leftarrow Y_{1:t}) \land (x^{Y'} \not\leftarrow Y'_{1:t}) \land (x^Y = x^{Y'})$, i.e.
+$x$ is overwritten in both sequences at some point between $t+1$ and $L$ or $L'$, but by the same component state.
+Since the first two conditions are again independent we have
+
+$$P(\textrm{Case 2}) = P(x^Y \not\leftarrow Y_{1:t})P(x^{Y'} \not\leftarrow Y'_{1:t})P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$$
+
+$$= [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t]P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$$
+
+where we recall that $Y_{1:t} = Y'_{1:t}$. When $\gamma = 1$ this simplifies to
+
+$$P(\textrm{Case 2}) = \frac{L-t}{L+1}\frac{L'-t}{L'+1}P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$$
+
+To find $P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$ we note that if both $x$'s are overwritten, they can end up with the same component state if (1) $x^Y$ and $x^{Y'}$ are taken from the same symbol $i$, whose probability we write as $P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow Y_{1:t})$ or (2) $x^Y$ and $x^{Y'}$ are taken from different symbols but end up with the same component state by chance.
+That is,
+
+$$P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) =$$
+
+$$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) + [1 - P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow Y_{1:t})]\frac{1}{Z}.$$
+
+$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}})$ is the sum over $i \in \{1, ..., M\}$ of the probabilities that both $x^Y$ and $x^{Y'}$ were taken from $i$.
+These events are independent, so:
+
+$$P(x^Y, x^{Y'} \leftarrow i|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = P(x^Y \leftarrow i|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}})P(x^{Y'} \leftarrow i|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}})$$
+
+$$ = \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}.$$
+
+When $\gamma = 1$, this simplifies to
+
+$$\frac{n^i_v}{L-t}\frac{n^i_{v'}}{L'-t}$$.
+
+Thus
+
+$$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}.$$
+
+When $\gamma = 1$ this is
+
+$$P(x^Y, x^{Y'} \leftarrow same|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = \sum_i \frac{n^i_vn^i_{v'}}{(L-t)(L'-t)} = \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)}$$
+
+i.e.
+when $\gamma = 1$ the probability that $x^Y$ came from the same symbol, given that it was overwritten in the second part of both sequence constructions, is just the normalized dot product of the symbol-count vectors for the second parts of the two sequences.
+
+Thus
+
+$$P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = $$
+
+$$\sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} + \left(1 - \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}$$
+
+When $\gamma = 1$
+
+$$P(x^Y = x^{Y'}|x^Y, x^{Y'} \not\leftarrow X^{Y_{1:t}}) = \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}.$$
+
+Thus
+
+$$P(\textrm{Case 2}) = [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t] \times$$
+
+$$\left[
+\sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} + \left(1 - \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
+\right]$$
+
+When $\gamma = 1$:
+
+$$P(\textrm{Case 2}) = \frac{L-t}{L+1}\frac{L'-t}{L'+1}\left[\frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}\right].$$
+
+\textbf{Case 3}: $(x^Y \leftarrow Y_{1:t}) \land (x^{Y'} \not\leftarrow Y'_{1:t}) \land (x^Y = x^{Y'})$, i.e.
+$x$ is untouched in the first sequence but overwritten in the second, but happens to be overwritten by $x^{Y_{1:t}}$.
+Once again the first two events are independent, so
+
+$$P(\textrm{Case 3}) = (t+1)^\gamma f^{\gamma,L}_t[1-(t+1)^\gamma f^{\gamma,L'}_t]P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}).$$
+
+When $\gamma = 1$:
+
+$$P(\textrm{Case 3}) = \frac{t+1}{L+1}\frac{L'-t}{L'+1}P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}).$$
+
+By similar reasoning as in Case 2, the event $(x^Y = x^{Y'}|x^Y \leftarrow X^{Y_{1:t}}, x^{Y'} \not\leftarrow X^{Y'_{1:t}})$ can occur if (1) $x^Y$ and $x^{Y'}$ are taken from the same symbol $i$ or (2) $x^Y$ and $x^{Y'}$ are taken from different symbols but end up with the same component state.
+
+Following similar logic as in Case 2, except swapping $Y^{t+1:L}$ with $Y^{1:t}$ we thus have
+
+$$P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}) =$$
+
+$$\left[
+\sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} 
++ \left(1 - \sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
+\right]$$
+
+When $\gamma = 1$:
+
+$$P(x^Y = x^{Y'}|x^Y \leftarrow Y_{1:t}, x^{Y'} \not\leftarrow Y'_{1:t}) =
+\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}$$
+
+Thus,
+
+$$P(\textrm{Case 3}) = $$
+$$(t+1)^\gamma f^{\gamma,L}_t[1-(t+1)^\gamma f^{\gamma,L'}_t] \times 
+\left[
+\sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} 
++ \left(1 - \sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
+\right]$$
+
+And when $\gamma = 1$:
+
+$$P(\textrm{Case 3}) = \frac{t+1}{L+1}\frac{L'-t}{L'+1}\left[
+\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}
+\right].$$
+
+\textbf{Case 4}: $(x^Y \not \leftarrow Y_{1:t}) \land (x^{Y'} \leftarrow Y'_{1:t}) \land (x^Y = x^{Y'})$, i.e.
+$x$ is untouched in the second sequence but overwritten in the first, but happens to be overwritten by $x^{Y'_{1:t}}$.
+This is symmetric to Case 3, so
+
+$$P(\textrm{Case 4}) = $$
+$$[1 - (t+1)^\gamma f^{\gamma,L}_t](t+1)^\gamma f^{\gamma,L'}_t \times \left[
+\sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t} 
++ \left(1 - \sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t}\right)\frac{1}{Z}
+\right]$$
+
+And when $\gamma = 1$:
+
+$$P(\textrm{Case 4}) = \frac{L'-t}{L+1}\frac{t+1}{L'+1}\left[
+\frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} + \left(1 - \frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} \right)\frac{1}{Z}
+\right].$$
+
+\textbf{Sum over cases}
+
+$$P(x^Y = x^{Y'}) = P(\textrm{Case 1}) + P(\textrm{Case 2}) + P(\textrm{Case 3}) + P(\textrm{Case 4}) = $$
+
+$$
+(t+1)^\gamma f^{\gamma,L}_t(t+1)^\gamma f^{\gamma,L'}_t + [1 - (t+1)^\gamma f^{\gamma,L}_t][1 - (t+1)^\gamma f^{\gamma,L'}_t] \times$$
+
+$$\left[
+\sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} + \left(1 - \sum_i \frac{\sum\limits_{v^i} f^{\gamma,L}_{v^i}}{1 - (t+1)^\gamma f^{\gamma,L}_t} \frac{\sum\limits_{(v')^i} f^{\gamma,L'}_{(v')^i}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
+\right]$$
+
+$$+(t+1)^\gamma f^{\gamma,L}_t[1-(t+1)^\gamma f^{\gamma,L'}_t] \times 
+\left[
+\sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t} 
++ \left(1 - \sum_i \frac{\sum\limits_{u^i} f_{u^i}^{\gamma, t} \sum\limits_{(v')^i} f_{(v')^i}^{\gamma,L'}}{1 - (t+1)^\gamma f^{\gamma,L'}_t}\right)\frac{1}{Z}
+\right]
+$$
+
+$$+[1 - (t+1)^\gamma f^{\gamma,L}_t](t+1)^\gamma f^{\gamma,L'}_t \times \left[
+\sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t} 
++ \left(1 - \sum_i \frac{\sum\limits_{v^i} f_{v^i}^{\gamma,L}\sum\limits_{u^i} f_{u^i}^{\gamma, t}}{1 - (t+1)^\gamma f^{\gamma,L}_t}\right)\frac{1}{Z}
+\right]
+$$
+
+When $\gamma = 1$ this simplifies to
+
+$$P(x^Y = x^{Y'}) =  \frac{t+1}{L+1}\frac{t+1}{L'+1}$$
+
+$$
++ \frac{L-t}{L+1}\frac{L'-t}{L'+1}\left[\frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} + \left(1 - \frac{\mathbf{n}^T_v\mathbf{n}_{v'}}{(L-t)(L'-t)} \right)\frac{1}{Z}\right]
+$$
+
+$$+\frac{t+1}{L+1}\frac{L'-t}{L'+1}\left[
+\frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)}+ \left(1 - \frac{\mathbf{n}^T_{u}\mathbf{n}_{v'}}{(t+1)(L'-t)} \right)\frac{1}{Z}
+\right]$$
+$$
++\frac{L'-t}{L+1}\frac{t+1}{L'+1}\left[
+\frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} + \left(1 - \frac{\mathbf{n}^T_{v}\mathbf{n}_{u}}{(L-t)(t+1)} \right)\frac{1}{Z}
+\right].
+$$
+
+\end{appendices}
 

### 2019-11-14 ms_neurips_2019/ms_neurips_2019.tex
Update for submission to arXiv.

@@
 Inspired by chromosomal crossover, random HD vectors are recursively interwoven, with a fraction of one vector's components masked out and replaced by those from another using a context-dependent mask.
-Unlike most HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
+Unlike many HD computing schemes, "crossover" codes highly overlap with their base elements' and sub-structures' codes without sacrificing relational information, allowing fast element readout and decoding by greedy reconstruction.
 Crossover is mathematically tractable and has several properties desirable for robust, flexible representation.
@@
 HDC's advantages are (1) random HD vectors tend to have very low overlap, enabling large element codebooks, (2) there is no hard limit on data complexity, and (3) encoding is often analytically tractable.
-While most systems must of course learn some environment statistics, HDC provides a compelling blueprint for flexible representation.
+While most systems must of course undergo some amount of training, HDC provides a compelling blueprint for flexible representation.
 
 An ideal code should allow fast readout of both a composition's base elements and their relations, but most HDC schemes support one or the other.
-Sums of random HD vectors, for instance, have high overlap (low distance under most metrics) to their elements, enabling element readout via overlaps with the sum \cite{Bloom:1970, Anderson:1973, Plate:1994}; but cannot encode relations.
+Sums of random HD vectors, for instance, have high overlap (low distance under usual metrics) to their elements, enabling element readout via overlaps with the sum \cite{Bloom:1970, Anderson:1973, Plate:1994}; but cannot encode relations.
 HDC operations like circular convolution \cite{Plate:1995}, permutation \cite{Gayler:1998, Sahlgren:2008}, or matrix multiplication \cite{Gosmann:2019}, yield codes with cue-recoverable relations (cue A recovers B from ((A,B),(C,D))), but low overlap to the composition's base elements, precluding fast element readout without cues (but see \cite{Rachkovskij:2001}); this also conflicts with our desire that AB's code should resemble A's more than C's.
@@
 Precisely, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q \equiv P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$.
-As exemplified by Figure 1c, crossover code components are uniform over $G$, i.i.d.
+As exemplified by Figure 1c, crossover code component distributions are uniform and i.i.d.
 for a given $Y$, and invariant to sequence statistics (e.g.
-length).
-The proof follows from the i.i.d.
+length); this results from the i.i.d.
 nature of $C, \mu$ and the interchangeability of component states, suggesting crossover as a robust, scale-invariant encoding.
@@
 (c) Decoding of *CAFE vs number of corrupted components at each encoding step for i.i.d.
-and correlated $\mu$.
+vs correlated $\mu$.
 (d) Decoding of *CICADA for i.i.d.
@@
 This is fixed by introducing correlations into $\mu_j(X, X')$ (i.e.
-making $\mu$ a smooth function of $X, X'$ [many methods work]).
+making $\mu$ a smooth function of $X, X'$ [see Appendix C]).
 The cost of this is decreased decoding accuracy for certain sequences, since correlated are less distinct than i.i.d.
@@
 
-While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built-in \cite{Zador:2019}.
+While intelligent systems must typically undergo at least some amount of training to function, it can be useful for certain computational features to be built in \cite{Zador:2019}.
 Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with the isolated scene elements.
 While many trained systems, e.g.
-translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}).
+translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016}, although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}.
 Alternatively, one can design HDC algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems.
 Basing training on convolutional binding, for instance, allowed an artificial network to efficiently and scalably learn knowledge graphs and out-perform state-of-the-art methods at link prediction \cite{Nickel:2016}.
-It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, as well as to guiding investigation of neural computation in biology.
+It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, and potentially guiding investigation of neural computation in biology.
 
@@
 This is relevant to biological implementation, as it allows one to envision $x_j$ not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank.
-For instance, each component of a crossover code could correspond to the attractor state of a small Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree.
-This could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code-overlap statistics to be usefully read out and manipulated by further downstream and recurrent computations.
+For instance, each component of a crossover code could correspond to the attractor state of a small Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a complex data object.
+This could serve, e.g., as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code-overlap statistics to be usefully read out and manipulated by further downstream and recurrent computations.
 
@@
 
-I acknowledge...
+I would like to thank Taliesin Beynon for many helpful suggestions in contextualizing this work within the broader scope of high-dimensional computing and vector symbolic architectures.
 
@@
 
-\section{Symbol/sequence code overlap derivations}
+\section{Code overlap derivations for symbols and sequences}
 
-Since $x^Y_j$ are i.i.d., the similarity or overlap between two sequence codes  (1 minus the Hamming distance) $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$.
+Since $x^Y_j$ are i.i.d., the similarity, or overlap, between two sequence codes  (1 minus the Hamming distance) $s(X^Y, X^{Y'}) \sim \frac{1}{N}\textrm{Bi}(N, q)$, where $\textrm{Bi}(N, q)$ is a binomial distribution with success probability $q = P(x^Y_j = x^{Y'}_j)$.
 The randomness giving rise to the distribution is over samples of $C$ and $\mu$.
@@
 
-Note that this is not a function of $x^Y$. Also note that $f^{1,L}_t = 1/(L+1)$, which is not a function of $t$.
+Note that this is not a function of $x^Y$. Also note that $f^{1,L}_t = 1/(L+1)$, which is independent of $t$.
 
@@
 < t_{n^i}$, with $\sum_i n^i = L$.
-We can divide the combinatorial ways $x^Y$ could have ended up being taken from $i$ into $n^i$ cases.
+We can divide the combinatorial ways $x^Y$ could have been taken from $i$ into $n^i$ cases.
 First, $x^Y$ could have been taken from $y_{t_1}$ then never overwritten, whose probability is $f^{\gamma,L}_{t_1}$.
@@
 
-\subsection{Derivation: Similarity between symbol and sequence codes}
+\subsection{Similarity between symbol and sequence codes}
 
@@
 
-\subsection{Derivation: Similarity between codes for two sequences}
+\subsection{Similarity between codes for two sequences}
 
@@
 
+\section{Correlated mask function construction}
+
+We introduced correlations into the mask function $\mu$ (Figure 2d) as follows.
+We first created a random, sparse, binary matrix $W$ of size $N \times 2N$ and with density $r$ (0.004 in Fig 2d), and assigned our component states $1, ..., Z$ to evenly spaced angles between $0$ and $2\pi$.
+The mask value $\mu_j(X, X')$ was then determined by (1) multiplying $W$ by the concatenated $2N$-dimensional vector $[X; X']$, (2) taking the circular mean $\phi$ of the result, and (3) letting $\mu_j(X, X')$ be equal to absolute angular distance between $\phi$ and $\pi$, divided by $\pi$.
+This led to mask function values uniformly distributed between $0$ and $1$ yet which were determined in a smoother way by the values of $X$ and $X'$ than when $\mu$ was sampled i.i.d.
+Upon quadrupling $N$ (Figure 2d magenta) we equivalently decreased $r$ by a factor of 4.
+
+\section{Code}
+
+All code used in the preparation of this manuscript was written in the Python programming language and is available at \url{https://github.com/rkp8000/crossing_over}.
+
 \end{appendices}
